### 2024-11-23 paper/paper.qmd
Initial upload of files

---
title: "Number of COVID-19 Cases Around the World"
subtitle: "Investigating the number of COVID-19 Cases, Tests, and Deaths by Country, Region, Age, and Sex"
author: 
  - Bruce Zhang
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)

# Load the dataset
analysis_data <- read.csv(here("data/02-analysis_data/analysis_data.csv"))

```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR] to analyze the data and to create graphs and models. Th packages that were used include tidyverse [@tidyverse], tidyr [@tidyr], dplyr [@dplyr], and rstanarm [@rstanarm]. Our data [@covid_data] was obtained from COVerAGE-DB. The data anlysis was conducted based on the guidance of @tellingstories.

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: cases-by-country
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary library
library(ggplot2)

# Create a bar plot of total cases per country
ggplot(analysis_data, aes(x = Country, y = Cases, fill = Country)) +
  geom_bar(stat = "identity", color = "lightblue", alpha = 0.
)

```

```{r}
#| echo: false
#| eval: true
#| label: cases-over-time
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)
library(lubridate) # For date handling

# Filter data to include all regions
filtered_data <- analysis_data %>%
  filter(Region == "All") %>%
  mutate(
    # Ensure the Date column is in proper Date format
    Date = as.Date(Date, format = "%Y-%m-%d"),
    # Extract the year and month to group by
    Month = floor_date(Date, unit = "month")
  )

# Create the scatterplot with aggregated monthly cases
ggplot(filtered_data, aes(x = Date, y = Cases, color = Country, group = Country)) +
  geom_point(alpha = 0.7) +   # Add scatter points
  geom_line(size = 1) +       # Add lines connecting points for each country
  labs(
    title = "Monthly COVID Cases in Canada (Urban Region)",
    x = "Month",
    y = "Total Number of Cases",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )

```

```{r}
#| echo: false
#| eval: true
#| label: cases-by-age
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
filtered_data <- analysis_data %>%
  filter(Country %in% selected_countries) %>%
  mutate(
    # Create age intervals
    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE)
  )

# Aggregate the number of cases by country and age interval
age_case_summary <- filtered_data %>%
  group_by(Country, Age_Interval) %>%
  summarise(Total_Cases = sum(Cases, na.rm = TRUE)) %>%
  ungroup()

# Create the faceted histogram
ggplot(age_case_summary, aes(x = Age_Interval, y = Total_Cases, fill = Country)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
  facet_wrap(~Country, scales = "free_y") + # Create a grid of histograms for each country
  labs(
    title = "COVID Cases by Age Interval in Selected Countries",
    x = "Age Interval",
    y = "Number of Cases",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "grey80"),
    legend.position = "none"
  )

```

```{r}
#| echo: false
#| eval: true
#| label: deaths-by-age
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
filtered_data <- analysis_data %>%
  filter(Country %in% selected_countries) %>%
  mutate(
    # Create age intervals
    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE)
  )

# Aggregate the number of deaths by country and age interval
age_death_summary <- filtered_data %>%
  group_by(Country, Age_Interval) %>%
  summarise(Total_Deaths = sum(Deaths, na.rm = TRUE)) %>% # Use Deaths column
  ungroup()

# Create the faceted histogram for deaths
ggplot(age_death_summary, aes(x = Age_Interval, y = Total_Deaths, fill = Country)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
  facet_wrap(~Country, scales = "free_y") + # Create a grid of histograms for each country
  labs(
    title = "COVID Deaths by Age Interval in Selected Countries",
    x = "Age Interval",
    y = "Number of Deaths",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "grey80"),
    legend.position = "none"
  )

```

```{r}
#| echo: false
#| eval: true
#| label: cases-by-sex
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries and group by gender and country
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
gender_case_summary <- analysis_data %>%
  filter(Country %in% selected_countries, Sex %in% c("f", "m")) %>% # Filter for selected countries and genders
  group_by(Country, Sex) %>%
  summarise(Total_Cases = sum(Cases, na.rm = TRUE)) %>% # Summarize total cases by gender and country
  ungroup()

# Create the faceted bar graph
ggplot(gender_case_summary, aes(x = Sex, y = Total_Cases, fill = Sex)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
  facet_wrap(~Country, scales = "free_y") + # Facet by country
  scale_fill_manual(values = c("f" = "pink", "m" = "blue")) + # Optional: Gender-specific colors
  labs(
    title = "COVID Cases by Gender in Selected Countries",
    x = "Gender",
    y = "Number of Cases",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    panel.spacing = unit(1, "lines") # Increase spacing between facets
  )

```

```{r}
#| echo: false
#| eval: true
#| label: deaths-by-sex
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries and group by gender and country
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
gender_death_summary <- analysis_data %>%
  filter(Country %in% selected_countries, Sex %in% c("f", "m")) %>% # Filter for selected countries and genders
  group_by(Country, Sex) %>%
  summarise(Total_Deaths = sum(Deaths, na.rm = TRUE)) %>% # Summarize total deaths by gender and country
  ungroup()

# Create the faceted bar graph
ggplot(gender_death_summary, aes(x = Sex, y = Total_Deaths, fill = Sex)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
  facet_wrap(~Country, scales = "free_y") + # Facet by country
  scale_fill_manual(values = c("f" = "pink", "m" = "blue")) + # Optional: Gender-specific colors
  labs(
    title = "COVID Deaths by Gender in Selected Countries",
    x = "Gender",
    y = "Number of Deaths",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    panel.spacing = unit(1, "lines") # Increase spacing between facets
  )

```

```{r}
#| echo: false
#| eval: true
#| label: cases-by-region
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries and group by region and country
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
region_case_summary <- analysis_data %>%
  filter(Country %in% selected_countries) %>% # Filter for selected countries
  group_by(Country, Region) %>%
  summarise(Total_Cases = sum(Cases, na.rm = TRUE)) %>% # Summarize total cases by region and country
  ungroup()

# Create the faceted bar graph
ggplot(region_case_summary, aes(x = Region, y = Total_Cases, fill = Region)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
  facet_wrap(~Country, scales = "free_y") + # Facet by country
  labs(
    title = "COVID Cases by Region in Selected Countries",
    x = "Region",
    y = "Number of Cases",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Rotate x-axis labels for readability
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    panel.spacing = unit(1, "lines") # Increase spacing between facets
  )

```

```{r}
#| echo: false
#| eval: true
#| label: deaths-by-region
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries and group by region and country
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
region_death_summary <- analysis_data %>%
  filter(Country %in% selected_countries) %>% # Filter for selected countries
  group_by(Country, Region) %>%
  summarise(Total_Deaths = sum(Deaths, na.rm = TRUE)) %>% # Summarize total deaths by region and country
  ungroup()

# Create the faceted bar graph
ggplot(region_death_summary, aes(x = Region, y = Total_Deaths, fill = Region)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
  facet_wrap(~Country, scales = "free_y") + # Facet by country
  labs(
    title = "COVID Deaths by Region in Selected Countries",
    x = "Region",
    y = "Number of Deaths",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Rotate x-axis labels for readability
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    panel.spacing = unit(1, "lines") # Increase spacing between facets
  )

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

```{r}

```

## Diagnostics

```{r}

```

\newpage

# References

### 2024-11-23 scripts/00-simulate_data.R

@@ -1,0 +1,44 @@
+#### Preamble ####
+# Purpose: Simulates a dataset of COVID cases and deaths relative to demographics and geographic location. 
+# Author: Bruce Zhang
+# Date: 21 November 2024
+# Contact: brucejc.zhang@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed? Make sure you are in the `COVID_demographics` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+set.seed(21)
+
+# Define parameters
+countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
+regions <- c("All", "Urban", "Rural")
+codes <- c("AF", "BR", "CA", "DK", "EG")
+dates <- seq(as.Date("2022-01-01"), as.Date("2022-12-31"), by = "1 month")
+sexes <- c("b", "m", "f")
+ages <- seq(0, 100, by = 10)
+age_int <- 10
+
+# Generate the data
+simulated_data <- expand.grid(
+  Country = countries,
+  Region = regions,
+  Code = codes,
+  Date = dates,
+  Sex = sexes,
+  Age = ages,
+  AgeInt = age_int
+)
+
+# Add random values for Cases, Deaths, and Tests
+simulated_data$Cases <- round(runif(nrow(simulated_data), min = 0, max = 20000))
+simulated_data$Deaths <- round(runif(nrow(simulated_data), min = 0, max = 500))
+simulated_data$Tests <- round(runif(nrow(simulated_data), min = 100, max = 50000))
+
+# Display a sample of the simulated data
+head(simulated_data)
+
+#### Save data ####
+write_csv(simulated_data, "data/00-simulated_data/simulated_data.csv")

### 2024-11-23 scripts/01-test_simulated_data.R

@@ -1,0 +1,78 @@
+#### Preamble ####
+# Purpose: Tests the structure, validity, and any inconsistencies or missing data of the simulated COVID demographics dataset. 
+# Author: Bruce Zhang
+# Date: 21 November 2024
+# Contact: brucejc.zhang@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+  # - The `tidyverse` package must be installed and loaded
+  # - The `dplyr` package must be installed and loaded
+  # - 00-simulate_data.R must have been run
+# Any other information needed? Make sure you are in the `COVID_demographics` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+
+analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+
+# Load necessary library
+library(dplyr)
+
+# Test 1: Verify structure of the dataset
+str(simulated_data)
+
+# Test 2: Check for missing values in key columns
+missing_summary <- simulated_data %>%
+  summarise(
+    Missing_Country = sum(is.na(Country)),
+    Missing_Region = sum(is.na(Region)),
+    Missing_Code = sum(is.na(Code)),
+    Missing_Date = sum(is.na(Date)),
+    Missing_Sex = sum(is.na(Sex)),
+    Missing_Age = sum(is.na(Age)),
+    Missing_Cases = sum(is.na(Cases)),
+    Missing_Deaths = sum(is.na(Deaths)),
+    Missing_Tests = sum(is.na(Tests))
+  )
+print("Missing values summary:")
+print(missing_summary)
+
+# Test 3: Summary statistics for numeric columns
+numeric_summary <- simulated_data %>%
+  summarise(
+    Min_Cases = min(Cases, na.rm = TRUE),
+    Max_Cases = max(Cases, na.rm = TRUE),
+    Mean_Cases = mean(Cases, na.rm = TRUE),
+    Min_Deaths = min(Deaths, na.rm = TRUE),
+    Max_Deaths = max(Deaths, na.rm = TRUE),
+    Mean_Deaths = mean(Deaths, na.rm = TRUE),
+    Min_Tests = min(Tests, na.rm = TRUE),
+    Max_Tests = max(Tests, na.rm = TRUE),
+    Mean_Tests = mean(Tests, na.rm = TRUE)
+  )
+print("Summary statistics for numeric columns:")
+print(numeric_summary)
+
+# Test 4: Verify unique combinations of key columns
+unique_combinations <- simulated_data %>%
[…]

### 2024-11-23 scripts/02-download_data.R

@@ -1,0 +1,58 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+# Load required libraries
+library(httr)
+library(here)
+
+# Set up directories
+data_dir <- here("data")
+if (!dir.exists(data_dir)) dir.create(data_dir)
+
+# Define data source
+data_url <- "https://osf.io/43ucn/download"
+output_file <- file.path(data_dir, "raw_data.csv")
+
+# Download the data
+if (!file.exists(output_file)) {
+  tryCatch({
+    download.file(data_url, destfile = output_file, mode = "wb")
+    message("Data downloaded successfully.")
+  }, error = function(e) {
+    stop("Error in downloading the data: ", e$message)
+  })
+} else {
+  message("File already exists. Skipping download.")
+}
+
+# Verify the download
+if (file.exists(output_file) && file.info(output_file)$size > 0) {
+  message("File downloaded and verified.")
+} else {
+  stop("Failed to download or verify the file.")
+}
+
+# Log the download
+download_log <- file.path(data_dir, "download_log.txt")
+cat(
+  paste(Sys.time(), "- Downloaded:", output_file, "\n"),
+  file = download_log,
+  append = TRUE
+)
+
+
+#### Save data ####
+# [...UPDATE THIS...]
+# change the_raw_data to whatever name you assigned when you downloaded it.
+Data <- read.csv(output_file)
+
+write_csv(Data, "data/01-raw_data/raw_data.csv") 
+
+         

### 2024-11-23 scripts/03-clean_data.R

@@ -1,0 +1,77 @@
+#### Preamble ####
+# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 6 April 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+install.packages("tidyr")
+# Load necessary libraries
+library(dplyr)
+library(tidyr)
+library(tidyverse)
+
+# Step 1: Remove unnecessary columns (if applicable)
+data_cleaned <- simulated_data %>%
+  select(Country, Region, Date, Sex, Age, Cases, Deaths, Tests)
+
+# Step 2: Handle missing values
+# Replace missing numeric values with 0 (if appropriate)
+data_cleaned <- data_cleaned %>%
+  mutate(
+    Cases = ifelse(is.na(Cases), 0, Cases),
+    Deaths = ifelse(is.na(Deaths), 0, Deaths),
+    Tests = ifelse(is.na(Tests), 0, Tests)
+  )
+
+# Step 3: Format dates
+data_cleaned <- data_cleaned %>%
+  mutate(
+    # Ensure Date is in proper date format
+    Date = as.Date(Date, format = "%d.%m.%Y"),
+    
+    # Extract year as a numeric variable
+    Year = as.numeric(format(Date, "%Y")),
+    
+    # Convert Date into a continuous variable (number of days since 1970-01-01)
+    Date_Continuous = as.numeric(Date)
+  )
+
+
+# Step 4: Ensure consistent data types
+data_cleaned <- data_cleaned %>%
+  mutate(
+    Country = as.factor(Country),
+    Region = as.factor(Region),
+    Sex = as.factor(Sex),
+    Age = as.numeric(Age),
+    Cases = as.numeric(Cases),
+    Deaths = as.numeric(Deaths),
+    Tests = as.numeric(Tests)
+  )
+
+# Remove duplicate rows (if any)
+data_cleaned <- data_cleaned %>%
+  distinct()
+
+# Check for and handle outliers (optional)
+# For example, cap unrealistic values for 'Cases', 'Deaths', or 'Tests'
+data_cleaned <- data_cleaned %>%
+  mutate(
[…]

### 2024-11-23 scripts/06-model_data.R

@@ -1,0 +1,52 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read data ####
+analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+
+### Model data ####
+first_model <-
+  stan_glm(
+    formula = flying_time ~ length + width,
+    data = analysis_data,
+    family = gaussian(),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_aux = exponential(rate = 1, autoscale = TRUE),
+    seed = 853
+  )
+
+install.packages("forecast")
+library(forecast)
+# Aggregate data by date
+time_series_data <- analysis_data %>%
+  group_by(Date) %>%
+  summarise(Total_Cases = sum(Cases, na.rm = TRUE))
+
+# Create a time series object
+ts_data <- ts(time_series_data$Total_Cases, frequency = 365)
+
+# Fit an ARIMA model
+fit <- auto.arima(ts_data)
+forecast(fit, h = 30) # Forecast next 30 days
+
+
+
+#### Save model ####
+saveRDS(
+  first_model,
+  file = "models/first_model.rds"
+)
+
+

### 2024-11-23 scripts/03-clean_data.R
Updated analysis data testing

@@ -15,7 +15,7 @@ library(tidyverse)
 
 # Step 1: Remove unnecessary columns (if applicable)
 data_cleaned <- simulated_data %>%
-  select(Country, Region, Date, Sex, Age, Cases, Deaths, Tests)
+  select(Country, Region, Date, Sex, Age, Cases, Deaths)
 
 # Step 2: Handle missing values
 # Replace missing numeric values with 0 (if appropriate)
@@ -49,12 +49,13 @@ data_cleaned <- data_cleaned %>%
     Age = as.numeric(Age),
     Cases = as.numeric(Cases),
     Deaths = as.numeric(Deaths),
-    Tests = as.numeric(Tests)
   )
 
 # Remove duplicate rows (if any)
 data_cleaned <- data_cleaned %>%
-  distinct()
+  distinct() %>%
+  filter(Country == "Canada") %>%
+  filter(Region %in% c("Urban", "Rural"), Sex %in% c("m", "f"))
 
 # Check for and handle outliers (optional)
 # For example, cap unrealistic values for 'Cases', 'Deaths', or 'Tests'
@@ -64,6 +65,9 @@ data_cleaned <- data_cleaned %>%
     Deaths = ifelse(Deaths > 1000, 1000, Deaths),
     Tests = ifelse(Tests > 500000, 500000, Tests)
   )
+
+data_cleaned <- data_cleaned %>%
+  filter(Deaths <= Cases)
 
 # Add calculated columns (if needed)
 # Example: Case Fatality Rate (CFR)

### 2024-11-23 scripts/04-test_analysis_data.R

@@ -1,69 +1,120 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests the analysis data for any potential problems that may compromise the analysis. 
+# Author: Bruce Zhang
+# Date: 23 September 2024
+# Contact: brucejc.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+ # - 03-clean_data.R must have been run
+# Any other information needed? Make sure you are in the `COVID_demographics` rproj
 
 
-#### Workspace setup ####
-library(tidyverse)
-library(testthat)
+# Load necessary libraries
+library(dplyr)
+library(ggplot2)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+# Test the integrity of the analysis dataset
+message("Testing the analysis data...")
+
+# Step 1: Load the data
+# Replace 'analysis_data_path' with the actual path to your data file
+analysis_data_path <- "data/02-analysis_data/analysis_data.csv"
+analysis_data <- read.csv(analysis_data_path)
+
+# Step 2: Basic Structure and Preview
+message("Previewing the dataset...")
+print(head(analysis_data))
+print(str(analysis_data))
+
+# Step 3: Check for Missing Values
+message("Checking for missing values...")
+missing_summary <- analysis_data %>%
+  summarise(across(everything(), ~ sum(is.na(.)))) %>%
+  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Missing_Count")
+
+print(missing_summary)
+
+# Step 4: Check for Duplicates
+message("Checking for duplicate rows...")
+duplicates <- analysis_data %>%
+  duplicated() %>%
+  sum()
+message(paste("Number of duplicate rows:", duplicates))
+
+# Step 5: Summary Statistics
+message("Generating summary statistics...")
+summary_statistics <- analysis_data %>%
+  summarise(
+    Total_Regions = n_distinct(Region),
+    Total_Countries = n_distinct(Country),
+    Total_Sexes = n_distinct(Sex),
[…]

### 2024-11-23 paper/paper.qmd
Updated model set up of paper

@@
 ---
-title: "Number of COVID-19 Cases Around the World"
-subtitle: "Investigating the number of COVID-19 Cases, Tests, and Deaths by Country, Region, Age, and Sex"
+title: "Number of COVID-19 Cases in Canada By Region, Age, and Sex"
+subtitle: "Predicting the Factors that Influence the Number of COVID-19 Cases and Deaths"
 author: 
   - Bruce Zhang
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+thanks: "Code and data are available at: [https://github.com/brucejczhang/covid_data](https://github.com/brucejczhang/covid_data)."
 date: today
@@
 analysis_data <- read.csv(here("data/02-analysis_data/analysis_data.csv"))
-
 ```
@@
 
+# Load necessary libraries
+library(dplyr)
+library(knitr)
+
+# Calculate Case Fatality Rate (CFR) and summarize the data
+summary_stats <- analysis_data %>%
+  mutate(
+    CFR = Deaths / (Cases + 1e-6) * 100 # Calculate Case Fatality Rate (percentage)
+  ) %>%
+  group_by(Region, Sex) %>% # Group by Region and Sex
+  summarise(
+    Mean_Age = mean(Age, na.rm = TRUE),             # Average age
+    Total_Cases = sum(Cases, na.rm = TRUE),         # Total cases
+    Total_Deaths = sum(Deaths, na.rm = TRUE),       # Total deaths
+    Mean_CFR = mean(CFR, na.rm = TRUE)              # Average Case Fatality Rate
+  ) %>%
+  ungroup() # Remove grouping for display purposes
+
+# Display the summary table
+kable(summary_stats, digits = 2, col.names = c(
+  "Region", "Sex", "Mean Age", "Total Cases", "Total Deaths", "Mean CFR (%)"
+))
+
 ```
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+A logistic model has been selected as part of this analysis to examine the relationship between region, sex, age, and the number of cases and deaths of COVID-19.
 
+The logistic regression model predicts the probability of a region or individual being classified as "high risk" based on key predictors.
+In this case, "high risk" is defined as regions where the death rate (CFR) exceeds the median death rate across the dataset.
+
+The CFR is calculated as: 
 \begin{align}
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+\text{CFR} = \frac{\text{Deaths}}{\text{Cases} + 10^{-6}} \times 100
 \end{align}
 
+The \textbf{box-and-whisker plot} visualizes the distribution of predicted probabilities (\( p \)) across different \textbf{age intervals}.
+Each box represents the interquartile range (IQR) of predicted probabilities, with the line inside the box indicating the median predicted probability for that age interval.
+Whiskers extend to 1.5 times the IQR, and any points outside this range are considered outliers.
+This visualization highlights the variability of predicted probabilities within each age group.
+
+The logistic regression model equation is:
+\begin{align}
+\text{logit}(p) = \ln\left(\frac{p}{1 - p}\right) = \beta_0 + \beta_1 \cdot \text{Tests} + \beta_2 \cdot \text{Age Interval} + \epsilon
+\end{align}
+
+where:
+\begin{itemize}
+    \item $p$: Predicted probability of being high risk ($p = \text{Predicted Probability}$).
+    \item $\text{logit}(p)$: Log-odds of the high-risk classification.
+    \item $\beta_0$: Intercept of the model.
+    \item $\beta_1, \beta_2$: Coefficients for the predictors (e.g., number of tests, age interval).
+    \item $\epsilon$: Error term.
+\end{itemize}
+
+This model explains how testing rates and age intervals affect the likelihood of a region or demographic being classified as high risk.
+The box-and-whisker plot shows the variability of predicted probabilities across age groups, helping to identify specific age intervals with consistently higher or lower risks.
+
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
+```{r}
+# Load necessary libraries
+library(dplyr)
+library(ggplot2)
+set.seed(21)
+
+# Step 1: Filter for Urban and Rural Regions Only
+filtered_data <- analysis_data %>%
+  mutate(
+    Death_Rate = Deaths / (Cases + 1e-6), # Calculate death rate (avoid division by zero)
+    High_Risk = ifelse(Death_Rate > median(Death_Rate, na.rm = TRUE), 1, 0) # Classify high-risk regions
+  )
+
+# Step 2: Logistic Regression Model
+# Fit a logistic regression model
+logistic_model <- glm(
+  High_Risk ~ Tests + Region + Age,
+  family = binomial(),
+  data = filtered_data
+)
+
+# Summary of the model
+summary(logistic_model)
+
+# Step 3: Add Predictions to the Dataset
+filtered_data <- filtered_data %>%
+  mutate(Predicted_Probability = predict(logistic_model, type = "response"))
+
+# Step 4: Visualize the Results
+# Plot predicted probabilities by region
+ggplot(filtered_data, aes(x = Region, y = Predicted_Probability, fill = Region)) +
+  geom_boxplot(alpha = 0.6) +
+  labs(
+    title = "Predicted Probability of High Risk by Region (Urban vs Rural)",
+    x = "Region",
+    y = "Predicted Probability of High Risk"
+  ) +
+  theme_minimal()
+
+```
+
+```{r}
+# Load necessary libraries
+library(dplyr)
+library(ggplot2)
+set.seed(21)
+
+# Step 1: Filter for only 'f' and 'm', and define High Risk
+filtered_data <- analysis_data %>%
+  mutate(
+    Death_Rate = Deaths / (Cases + 1e-6), # Calculate death rate (avoid division by zero)
+    High_Risk = ifelse(Death_Rate > median(Death_Rate, na.rm = TRUE), 1, 0) # Classify high-risk regions
+  )
+
+# Step 2: Logistic Regression Model
+# Fit a logistic regression model
+logistic_model <- glm(
+  High_Risk ~ Tests + Sex + Age,
+  family = binomial(),
+  data = filtered_data
+)
+
+# Summary of the model
+summary(logistic_model)
+
+# Step 3: Add Predictions to the Dataset
+filtered_data <- filtered_data %>%
+  mutate(Predicted_Probability = predict(logistic_model, type = "response"))
+
+# Step 4: Visualize the Results
+# Plot predicted probabilities by sex
+ggplot(filtered_data, aes(x = Sex, y = Predicted_Probability, fill = Sex)) +
+  geom_boxplot(alpha = 0.6) +
+  labs(
+    title = "Predicted Probability of High Risk by Sex",
+    x = "Sex",
+    y = "Predicted Probability of High Risk"
+  ) +
+  theme_minimal()
+
+```
+
+```{r}
+# Load necessary libraries
+library(dplyr)
+library(ggplot2)
+set.seed(21)
+
+# Step 1: Create Age Intervals and Define High Risk
+filtered_data <- analysis_data %>%
+  mutate(
+    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE), # Create age intervals
+    Death_Rate = Deaths / (Cases + 1e-6), # Calculate death rate (avoid division by zero)
+    High_Risk = ifelse(Death_Rate > median(Death_Rate, na.rm = TRUE), 1, 0) # Classify high-risk regions
+  ) %>%
+  filter(!is.na(Age_Interval)) # Remove any rows with missing age intervals
+
+# Step 2: Logistic Regression Model
+# Fit a logistic regression model
+logistic_model <- glm(
+  High_Risk ~ Tests + Age_Interval,
+  family = binomial(),
+  data = filtered_data
+)
+
+# Summary of the model
+summary(logistic_model)
+
+# Step 3: Add Predictions to the Dataset
+filtered_data <- filtered_data %>%
+  mutate(Predicted_Probability = predict(logistic_model, type = "response"))
+
+# Step 4: Visualize the Results
+# Plot predicted probabilities by age interval
+ggplot(filtered_data, aes(x = Age_Interval, y = Predicted_Probability, fill = Age_Interval)) +
+  geom_boxplot(alpha = 0.6) +
+  labs(
+    title = "Predicted Probability of High Risk by Age Interval",
+    x = "Age Interval",
+    y = "Predicted Probability of High Risk"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for readability
+
+```
 
@@
 # Filter data for the selected 5 countries
-selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
+selected_countries <- c("Canada")
 filtered_data <- analysis_data %>%
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 

### 2024-11-25 paper/paper.qmd
Updated content and figures in paper

@@
 
-Overview paragraph
+The COVID-19 pandemic is something that is difficult to forget.
+Although it is now in the past, the effects that it had on society and how we function as individuals is still profound.
+At the same time, the statistical information that sprouted from the pandemic has a high value for analysis.
+This is not only to better understand the pandemic and how it affected populations but also to form a strong idea of its patterns to suggest tactics for dealing with pandemics shall something similar happen in the future.
 
-Estimand paragraph
+Studies have been done in recent years in attempts to utilize the data that was gathered during COVID times to better advise future preventions and treatment for pandemic situations.
+Many studies look into the relationship between demographic information of populations and their likelihood of getting diagnosed or fatality.
+One study found that communities on hillsides and other locations with enhanced ventilation and oxygen had reduced numbers of COVID cases [@cite-covid_demographics].
+Other studies have found relationships between exercise and cardiovascular function and the likelihood of getting COVID [@cite-cardio_and_covid].
+Although many relationships have been examined, there has been a lack of studies focusing on a broad set of demographic traits such as age and sex.
+There has also not been cross comparisons of region and demographic aspects.
+Another gap is the lack of country-specific analysis, particularly Canada.
 
-Results paragraph
+This analysis looks at COVID-19 data from 2022 in Canada and analyzes how the different population and individual-level variables, such as age, sex, and region of life, influenced the risk of being diagnosed with COVID and the risk of fatality.
+The paper uses the demographic data gathered from 2022 to predict the characteristics of populations and individuals that may be at higher risk of fatality once they have been diagnosed with COVID.
+This analysis may have broader implications for preventative measures for flues, other diseases, and pandemics in the future.
 
-Why it matters paragraph
+The analysis focuses on the outcome variable of predicted probability of high risk, which is a function of number of cases and number of deaths per datapoint and is further defined in @sec-model.
+The model predicts the probability of high risk as a result of a series of predictor variables including region, sex, and age of the reported data.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+Results paragraph (what was found?)
+
+Why it matters paragraph (why it matters?)
+
+The remainder of this paper is structured as follows.
+@sec-data highlights the characteristics of the dataset.
+The section summarizes the data through a series of summary statistics (@fig-summary) and represents the data in a visual way where specific trends and patterns can be observed (@fig-cases_over_time, @fig-cfr).
+@sec-model creates a logistic regression model predicting the probability of high risk depending on the region, sex, and age of the data point.
+The section includes model figures that display the predicted probabilities of high risk for the predictor variables respectively.
 
@@
 
-We use the statistical programming language R [@citeR] to analyze the data and to create graphs and models.
-Th packages that were used include tidyverse [@tidyverse], tidyr [@tidyr], dplyr [@dplyr], and rstanarm [@rstanarm].
-Our data [@covid_data] was obtained from COVerAGE-DB.
+I use the statistical programming language R [@citeR] to analyze the data and to create graphs and models.
+Th packages that were used include tidyverse [@tidyverse], tidyr [@tidyr], dplyr [@dplyr], caret [@caret], and pROC [@pROC].
+My data [@covid_data] was obtained from COVerAGE-DB, which was housed in Open Science Framework.
 The data anlysis was conducted based on the guidance of @tellingstories.
-
-Overview text
 
@@
 
+The dataset used in this analysis was gathered by COVerAGE-DB, a database that focuses on COVID related data.
+COVerAGE-DB often gathers data relating to COVID such as the number of cases, deaths, tests, and vaccinations through governmental institutes such as health ministries and statistical offices.
+This data is then organized by other variables such as sex, region, country, and age and presented to the general public.
+
 Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
 
-## Outcome variables
+## Data Cleaning
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+The original dataset contained data from five different countries including.
+For the purpose to focus on data in Canada for this analysis, the datapoints for other countries were removed.
+Further cleaning was done to remove any problematic datapoints such as when the number of deaths exceeded the number of cases.
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+## Outcome Variables and Predictor Variables
+
+The cleaned dataset contains variables including the country, region, date, sex, and age where the number of cases, deaths, and tests were reported.
+@fig-head_analysis_data shows a sample of the dataset, presenting the first six rows of the analysis data.
+The outcome variables highlighted include the total number of cases and the total number of deaths.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-head_analysis_data
+#| fig-cap: First six rows of the dataset after cleaning
 #| echo: false
+#| warning: false
+#| message: false
+
+# Load the necessary library
+library(knitr)
+
+# Display the first few rows of analysis_data
+head_table <- head(analysis_data)
+
+# Render the table with kable
+kable(
+  head_table,
+  booktabs = TRUE,  # Add better table styling
+)
+```
+
+```{r}
+#| label: fig-summary
+#| fig-cap: Summary Statistics of COVID-19 Cases in Canada
+#| echo: false
+#| warning: false
+#| message: false
 
@@
 
-# Calculate Case Fatality Rate (CFR) and summarize the data
-summary_stats <- analysis_data %>%
-  mutate(
-    CFR = Deaths / (Cases + 1e-6) * 100 # Calculate Case Fatality Rate (percentage)
+# Create the summary statistics table
+summary_table <- analysis_data %>%
+  filter(Region %in% c("Urban", "Rural"), Sex %in% c("f", "m")) %>%
+  group_by(Region, Sex) %>%
+  summarise(
+    Mean_Age = round(mean(Age, na.rm = TRUE), 2), # Rounded Mean Age
+    Total_Cases = sum(Cases, na.rm = TRUE),
+    Total_Deaths = sum(Deaths, na.rm = TRUE),
+    Average_Cases = round(mean(Cases, na.rm = TRUE), 2), # Rounded Average Cases
+    Average_Deaths = round(mean(Deaths, na.rm = TRUE), 2), # Rounded Average Deaths
+    CFR = round((Total_Deaths / Total_Cases) * 100, 2) # Rounded CFR (%)
   ) %>%
-  group_by(Region, Sex) %>% # Group by Region and Sex
-  summarise(
-    Mean_Age = mean(Age, na.rm = TRUE),             # Average age
-    Total_Cases = sum(Cases, na.rm = TRUE),         # Total cases
-    Total_Deaths = sum(Deaths, na.rm = TRUE),       # Total deaths
-    Mean_CFR = mean(CFR, na.rm = TRUE)              # Average Case Fatality Rate
-  ) %>%
-  ungroup() # Remove grouping for display purposes
+  ungroup()
 
-# Display the summary table
-kable(summary_stats, digits = 2, col.names = c(
-  "Region", "Sex", "Mean Age", "Total Cases", "Total Deaths", "Mean CFR (%)"
-))
+# Render the table using kable
+summary_table %>%
+  kable(
+    col.names = c("Region", "Sex", "Mean Age", "Total Cases", "Total Deaths", 
+                  "Average Cases", "Average Deaths", "CFR (%)")
+  )
 
@@
 
-Talk more about it.
+@fig-summary further breaks down the data summarizing the mean number of cases and deaths by region and sex.
+This creates an all-rounded summary of the dataset and gives a preliminary idea on the differences in cases and deaths depending on the categories of sex and region.
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+### Number of Cases and Number of Deaths
+
+The total and average numbers of cases and deaths are key outcome variables that can be correlated with the categorical predictors.
+These number can demonstrate the differences in COVID-19 susceptibility and threat level based on the characteristics of the region and the individual.
+
+### Case Fatality Rate
+
+The case fatality rate (CFR) is an additional aspect of the summary that was calculated to standardize the deaths and cases by category.
+This can be used as a processed outcome variable that can better represent the dataset and aid modeling later on.
+The total number of deaths and cases may lead to inaccurate representations of a region, sex, or age group due to the absolute number of individuals that fall within the category.
+The CFR value allows better comparison across categories.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| echo: false
+#| eval: true
+#| label: fig-cases_over_time
+#| fig-cap: "Number of cases per month in 2022 in Canada"
+#| warning: false
+
+# Load necessary libraries
+library(dplyr)
+library(ggplot2)
+library(lubridate)
+set.seed(21)
+
+# Ensure Date is a proper Date object
+analysis_data <- analysis_data %>%
+  mutate(Date = as.Date(Date))
+
+# Aggregate data by month and year
+aggregated_data <- analysis_data %>%
+  mutate(Month_Year = floor_date(Date, unit = "month")) %>% # Group by month
+  group_by(Month_Year) %>%
+  summarise(
+    Total_Cases = sum(Cases, na.rm = TRUE) # Summing cases for each month
+  ) %>%
+  ungroup()
+
+# Create the bar graph with trend line
+ggplot(aggregated_data, aes(x = Month_Year, y = Total_Cases)) +
+  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) + # Bar plot
+  geom_smooth(method = "loess", se = TRUE, color = "red", size = 1.2) + # Trend line
+  labs(
+    title = "",
+    x = "Month-Year",
+    y = "Total Cases"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels
+    axis.title = element_text(size = 14),
+    plot.title = element_text(size = 16, face = "bold")
+  )
+
+```
+
+### Time
+
+The time of year can also be correlated with the likelihood of getting COVID-19 independent of other predictor variables.
+As shown in @fig-cases_over_time, there is a weak pattern of increased total number of cases in the early summer months of May and June.
+The number of cases in winter and near winter months, specifically January and October, seem to be lower.
+
+```{r}
+#| label: fig-cfr
+#| fig-cap: Case Fatality Rate (CFR) by Age Group, Region, and Sex
 #| echo: false
@@
 
+# Load necessary libraries
+library(ggplot2)
+library(dplyr)
+set.seed(21)
+
+# Filter data to include only Urban and Rural regions and relevant sexes
+filtered_data <- analysis_data %>%
+  filter(Region %in% c("Urban", "Rural"), Sex %in% c("f", "m")) %>% # Filter for Urban, Rural, and f/m genders
+  mutate(Age_Group = cut(Age, breaks = seq(0, 100, by = 10), include.lowest = TRUE)) # Create age intervals
+
+# Aggregate data by Age_Group, Region, and Sex, and calculate CFR
+aggregated_data <- filtered_data %>%
+  group_by(Age_Group, Region, Sex) %>%
+  summarise(
+    Total_Deaths = sum(Deaths, na.rm = TRUE),  # Sum up deaths
+    Total_Cases = sum(Cases, na.rm = TRUE),    # Sum up cases
+    CFR = (Total_Deaths / Total_Cases) * 100   # Calculate CFR
+  ) %>%
+  ungroup()
+
+# Create the bar graph for CFR by age group, region, and sex
+ggplot(aggregated_data, aes(x = Age_Group, y = CFR, fill = Sex)) +
+  geom_bar(stat = "identity", position = "dodge") + # Dodge bars to show sexes side by side
+  facet_wrap(~Region, ncol = 1) + # Separate plots for Urban and Rural regions
+  labs(
+    title = "",
+    x = "Age Group",
+    y = "CFR (%)",
+    fill = "Sex"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels for readability
+    axis.title = element_text(size = 14),
+    plot.title = element_text(size = 16, face = "bold"),
+    strip.text = element_text(size = 12, face = "bold") # Adjust facet labels
+  )
+
+```
+### Age, Sex, and Region
+
+Age, sex, and region are key predictor variables that are examined in this analysis.
+In @fig-cfr, these variables are organized relative to the CFR, which calculates a rate based on the number of cases and number of deaths and allows standardized comparisons across different predictor variables.
+@fig-cfr shows that the urban and rural regions have similar patterns in CFR across different age groups.
+The CFR for males tend to be higher more often than not compared to the CFR of females, indicating that males may have a higher likelihood of death after diagnosis of COVID in Canada.
+The pattern of CFR in relation to age seem to vary quite significantly between urban and rural regions and between male and females.
+For females, the highest CFR seems to be for individuals aged under 30 and for those around 70 to 80 regardless of region.
+Male CFR values peak at age interval 20 to 30 for urban regions and 30 to 40 for rural regions.
+
+@fig-cfr_by_region shows a more direct comparison of the CFR between urban and rural regions of Canada, combining different age groups and sexes.
+This figure shows that the CFR in rural regions is noticeably higher than that of urban.
+Although the difference is small, it is still a considerable size difference when considering the number of people that a small percentage can be responsible for, based on the summary statistics from @fig-summary.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-cfr_by_region
+#| fig-cap: "Case Fatality Rate (CFR) by Region in 2022 in Canada"
+#| warning: false
+
+# Load necessary libraries
+library(ggplot2)
+library(dplyr)
+set.seed(21)
+
+# Filter data for the selected 5 countries and group by region and country
+selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
+region_cfr_summary <- analysis_data %>%
+  filter(Country %in% selected_countries, Region %in% c("Urban", "Rural")) %>% # Filter for selected countries and regions
+  group_by(Country, Region) %>%
+  summarise(
+    Total_Deaths = sum(Deaths, na.rm = TRUE),  # Summarize total deaths
+    Total_Cases = sum(Cases, na.rm = TRUE),    # Summarize total cases
+    CFR = (Total_Deaths / Total_Cases) * 100  # Calculate CFR
+  ) %>%
+  ungroup()
+
+# Create the faceted bar graph for CFR by region
+ggplot(region_cfr_summary, aes(x = Region, y = CFR, fill = Region)) +
+  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with CFR
+  scale_fill_manual(values = c("Urban" = "steelblue", "Rural" = "forestgreen")) + # Optional: Region-specific colors
+  labs(
+    title = "",
+    x = "Region",
+    y = "CFR (%)",
+    fill = "Region"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(size = 12),
+    axis.title = element_text(size = 14),
+    strip.text = element_text(size = 14),
+    panel.spacing = unit(1, "lines") # Increase spacing between facets
+  )
+
 ```
 
-Talk way more about it.
-
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
-
-# Model
+# Model {#sec-model}
 
@@
 
+### Model justification
+
+We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+
+We can use maths by including latex between dollar signs, for instance $\theta$.
+
+# Results {#sec-results}
+
+Our results are summarized in @tbl-modelresults.
+
 ```{r}
+#| label: model-risk_urban_vs_rural
+#| fig-cap: Predicted probability of high risk of COVID-19 by region (Urban vs. Rural)
+#| echo: false
+#| warning: false
+#| message: false
+
 # Load necessary libraries
@@
   labs(
-    title = "Predicted Probability of High Risk by Region (Urban vs Rural)",
+    title = "",
     x = "Region",
@@
 ```{r}
+#| label: model-risk_sex
+#| fig-cap: Predicted probability of high risk based on sex
+#| echo: false
+#| warning: false
+#| message: false
+
 # Load necessary libraries
@@
   labs(
-    title = "Predicted Probability of High Risk by Sex",
+    title = "",
     x = "Sex",
@@
 ```{r}
+#| label: model-risk_age
+#| fig-cap: Predicted probability of high risk based on age intervals
+#| echo: false
+#| warning: false
+#| message: false
+
 # Load necessary libraries
@@
   labs(
-    title = "Predicted Probability of High Risk by Age Interval",
+    title = "",
     x = "Age Interval",
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for readability
-
-```
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: cases-by-country
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-# Load necessary library
-library(ggplot2)
-
-# Create a bar plot of total cases per country
-ggplot(analysis_data, aes(x = Country, y = Cases, fill = Country)) +
-  geom_bar(stat = "identity", color = "lightblue", alpha = 0.
-)
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: cases-over-time
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-# Load necessary libraries
-library(ggplot2)
-library(dplyr)
-library(lubridate) # For date handling
-
-# Filter data to include all regions
-filtered_data <- analysis_data %>%
-  filter(Region == "All") %>%
-  mutate(
-    # Ensure the Date column is in proper Date format
-    Date = as.Date(Date, format = "%Y-%m-%d"),
-    # Extract the year and month to group by
-    Month = floor_date(Date, unit = "month")
-  )
-
-# Create the scatterplot with aggregated monthly cases
-ggplot(filtered_data, aes(x = Date, y = Cases, color = Country, group = Country)) +
-  geom_point(alpha = 0.7) +   # Add scatter points
-  geom_line(size = 1) +       # Add lines connecting points for each country
-  labs(
-    title = "Monthly COVID Cases in Canada (Urban Region)",
-    x = "Month",
-    y = "Total Number of Cases",
-    color = "Country"
-  ) +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    panel.grid.minor = element_blank()
-  )
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: cases-by-age
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-# Load necessary libraries
-library(ggplot2)
-library(dplyr)
-
-# Filter data for the selected 5 countries
-selected_countries <- c("Canada")
-filtered_data <- analysis_data %>%
-  filter(Country %in% selected_countries) %>%
-  mutate(
-    # Create age intervals
-    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE)
-  )
-
-# Aggregate the number of cases by country and age interval
-age_case_summary <- filtered_data %>%
-  group_by(Country, Age_Interval) %>%
-  summarise(Total_Cases = sum(Cases, na.rm = TRUE)) %>%
-  ungroup()
-
-# Create the faceted histogram
-ggplot(age_case_summary, aes(x = Age_Interval, y = Total_Cases, fill = Country)) +
-  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
-  facet_wrap(~Country, scales = "free_y") + # Create a grid of histograms for each country
-  labs(
-    title = "COVID Cases by Age Interval in Selected Countries",
-    x = "Age Interval",
-    y = "Number of Cases",
-    fill = "Country"
-  ) +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    panel.grid.major = element_line(color = "grey80"),
-    legend.position = "none"
-  )
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: deaths-by-age
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-# Load necessary libraries
-library(ggplot2)
-library(dplyr)
-
-# Filter data for the selected 5 countries
-selected_countries <- c("Canada")
-filtered_data <- analysis_data %>%
-  filter(Country %in% selected_countries) %>%
-  mutate(
-    # Create age intervals
-    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE)
-  )
-
-# Aggregate the number of deaths by country and age interval
-age_death_summary <- filtered_data %>%
-  group_by(Country, Age_Interval) %>%
-  summarise(Total_Deaths = sum(Deaths, na.rm = TRUE)) %>% # Use Deaths column
-  ungroup()
-
-# Create the faceted histogram for deaths
-ggplot(age_death_summary, aes(x = Age_Interval, y = Total_Deaths, fill = Country)) +
-  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
-  facet_wrap(~Country, scales = "free_y") + # Create a grid of histograms for each country
-  labs(
-    title = "COVID Deaths by Age Interval in Selected Countries",
-    x = "Age Interval",
-    y = "Number of Deaths",
-    fill = "Country"
-  ) +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    panel.grid.major = element_line(color = "grey80"),
-    legend.position = "none"
-  )
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: cases-by-sex
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-# Load necessary libraries
-library(ggplot2)
-library(dplyr)
-
-# Filter data for the selected 5 countries and group by gender and country
-selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
-gender_case_summary <- analysis_data %>%
-  filter(Country %in% selected_countries, Sex %in% c("f", "m")) %>% # Filter for selected countries and genders
-  group_by(Country, Sex) %>%
-  summarise(Total_Cases = sum(Cases, na.rm = TRUE)) %>% # Summarize total cases by gender and country
-  ungroup()
-
-# Create the faceted bar graph
-ggplot(gender_case_summary, aes(x = Sex, y = Total_Cases, fill = Sex)) +
-  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
-  facet_wrap(~Country, scales = "free_y") + # Facet by country
-  scale_fill_manual(values = c("f" = "pink", "m" = "blue")) + # Optional: Gender-specific colors
-  labs(
-    title = "COVID Cases by Gender in Selected Countries",
-    x = "Gender",
-    y = "Number of Cases",
-    fill = "Gender"
-  ) +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(size = 12),
-    axis.title = element_text(size = 14),
-    strip.text = element_text(size = 14),
-    panel.spacing = unit(1, "lines") # Increase spacing between facets
-  )
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: deaths-by-sex
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-# Load necessary libraries
-library(ggplot2)
-library(dplyr)
-
-# Filter data for the selected 5 countries and group by gender and country
-selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
-gender_death_summary <- analysis_data %>%
-  filter(Country %in% selected_countries, Sex %in% c("f", "m")) %>% # Filter for selected countries and genders
-  group_by(Country, Sex) %>%
-  summarise(Total_Deaths = sum(Deaths, na.rm = TRUE)) %>% # Summarize total deaths by gender and country
-  ungroup()
-
-# Create the faceted bar graph
-ggplot(gender_death_summary, aes(x = Sex, y = Total_Deaths, fill = Sex)) +
-  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
-  facet_wrap(~Country, scales = "free_y") + # Facet by country
-  scale_fill_manual(values = c("f" = "pink", "m" = "blue")) + # Optional: Gender-specific colors
-  labs(
-    title = "COVID Deaths by Gender in Selected Countries",
-    x = "Gender",
-    y = "Number of Deaths",
-    fill = "Gender"
-  ) +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(size = 12),
-    axis.title = element_text(size = 14),
-    strip.text = element_text(size = 14),
-    panel.spacing = unit(1, "lines") # Increase spacing between facets
-  )
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: cases-by-region
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-# Load necessary libraries
-library(ggplot2)
-library(dplyr)
-
-# Filter data for the selected 5 countries and group by region and country
-selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
-region_case_summary <- analysis_data %>%
-  filter(Country %in% selected_countries) %>% # Filter for selected countries
-  group_by(Country, Region) %>%
-  summarise(Total_Cases = sum(Cases, na.rm = TRUE)) %>% # Summarize total cases by region and country
-  ungroup()
-
-# Create the faceted bar graph
-ggplot(region_case_summary, aes(x = Region, y = Total_Cases, fill = Region)) +
-  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
-  facet_wrap(~Country, scales = "free_y") + # Facet by country
-  labs(
-    title = "COVID Cases by Region in Selected Countries",
-    x = "Region",
-    y = "Number of Cases",
-    fill = "Region"
-  ) +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Rotate x-axis labels for readability
-    axis.title = element_text(size = 14),
-    strip.text = element_text(size = 14),
-    panel.spacing = unit(1, "lines") # Increase spacing between facets
-  )
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: deaths-by-region
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-# Load necessary libraries
-library(ggplot2)
-library(dplyr)
-
-# Filter data for the selected 5 countries and group by region and country
-selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
-region_death_summary <- analysis_data %>%
-  filter(Country %in% selected_countries) %>% # Filter for selected countries
-  group_by(Country, Region) %>%
-  summarise(Total_Deaths = sum(Deaths, na.rm = TRUE)) %>% # Summarize total deaths by region and country
-  ungroup()
-
-# Create the faceted bar graph
-ggplot(region_death_summary, aes(x = Region, y = Total_Deaths, fill = Region)) +
-  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
-  facet_wrap(~Country, scales = "free_y") + # Facet by country
-  labs(
-    title = "COVID Deaths by Region in Selected Countries",
-    x = "Region",
-    y = "Number of Deaths",
-    fill = "Region"
-  ) +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Rotate x-axis labels for readability
-    axis.title = element_text(size = 14),
-    strip.text = element_text(size = 14),
-    panel.spacing = unit(1, "lines") # Increase spacing between facets
-  )
 

### 2024-11-25 scripts/03-clean_data.R

@@ -15,7 +15,7 @@ library(tidyverse)
 
 # Step 1: Remove unnecessary columns (if applicable)
 data_cleaned <- simulated_data %>%
-  select(Country, Region, Date, Sex, Age, Cases, Deaths)
+  select(Country, Region, Date, Sex, Age, Cases, Deaths, Tests)
 
 # Step 2: Handle missing values
 # Replace missing numeric values with 0 (if appropriate)
@@ -31,12 +31,6 @@ data_cleaned <- data_cleaned %>%
   mutate(
     # Ensure Date is in proper date format
     Date = as.Date(Date, format = "%d.%m.%Y"),
-    
-    # Extract year as a numeric variable
-    Year = as.numeric(format(Date, "%Y")),
-    
-    # Convert Date into a continuous variable (number of days since 1970-01-01)
-    Date_Continuous = as.numeric(Date)
   )
 
 

### 2024-11-25 scripts/06-model_data.R

@@ -7,46 +7,199 @@
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
+# Install and load necessary pacakges/libraries
+install.packages("caret")
+install.packages("pROC")
+library(ggplot2)
+library(dplyr)
+library(caret)
+library(pROC)
 
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
-
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
-
-install.packages("forecast")
-library(forecast)
-# Aggregate data by date
-time_series_data <- analysis_data %>%
-  group_by(Date) %>%
-  summarise(Total_Cases = sum(Cases, na.rm = TRUE))
-
-# Create a time series object
-ts_data <- ts(time_series_data$Total_Cases, frequency = 365)
-
-# Fit an ARIMA model
-fit <- auto.arima(ts_data)
-forecast(fit, h = 30) # Forecast next 30 days
-
-
-
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+# Building a logistic regression model for predicted high risk and age
+# Step 1: Build the logistic regression model
+logistic_model <- glm(
+  High_Risk ~ Age + Region + Sex,  # Replace with relevant predictors
+  data = analysis_data,
+  family = binomial()
 )
 
+# Step 2: Model Summary and Diagnostics
+cat("Model Summary:\n")
+print(summary(logistic_model))
+cat("\nAIC:", AIC(logistic_model), "\nBIC:", BIC(logistic_model), "\n")
 
+# Step 3: Predicted Probabilities and Classification
+analysis_data <- analysis_data %>%
+  mutate(
+    Predicted_Probability = predict(logistic_model, type = "response"),
[…]

### 2024-11-25 scripts/02-download_data.R
Updated preamble of scripts

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from Open Science Framework (OSF)
+# Author: Bruce Zhang
+# Date: 25 November 2024
+# Contact: brucejc.zhang@outlook.com
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+# Any other information needed? Make sure you are in the `COVID_demographics` rproj
 
 
 # Load required libraries

### 2024-11-25 scripts/03-clean_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw data on COVID-19 to only contain data for Canada and removes unecessary data
+# Author: Bruce Zhang
+# Date: 25 November 2024
+# Contact: brucejc.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+# Any other information needed? Make sure you are in the `COVID_demographics` rproj
 
 install.packages("tidyr")
 # Load necessary libraries

### 2024-11-25 scripts/06-model_data.R

@@ -1,11 +1,12 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models the predicted probability of high risk based on the predictor variables of the dataset
+# Author: Bruce Zhang
+# Date: 25 November 2024
+# Contact: brucejc.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+  # - 03-clean_data.R must have been run
+# Any other information needed? Make sure you are in the `COVID_demographics` rproj
 
 # Install and load necessary pacakges/libraries
 install.packages("caret")

### 2024-11-26 other/datasheet/datasheet_covid_data.qmd
Updated paper, models, and citations

---
title: "Datasheet for 'COVerAGE-DB' COVID-19 Dataset"
subtitle: "A summary of characteristics of the data and how it was used"
author: 
  - Bruce Zhang
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to analyze COVID-19 trends across different regions, sexes, and age groups. It allows for investigating relationships between predictors like tests, deaths, cases, and computed metrics like case fatality rates (CFR). The goal was to enable public health insights and predictive modeling.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset originates from publicly available sources compiled and structured by organizations such as the COVerAGE-DB project, which aggregates global COVID-19 data by demographic breakdowns.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - Funding sources include research initiatives and public health institutions supporting COVID-19 data dissemination, though specifics depend on the data contributors (e.g., COVerAGE-DB).
4. *Any other comments?*
    - TBD

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Instances represent demographic and epidemiological data for COVID-19, including age, region, sex, number of tests, cases, deaths, and derived metrics (e.g., CFR).
2. *How many instances are there in total (of each type, if appropriate)?*
	- The dataset contains thousands of rows representing data aggregated at various levels (e.g., country-region-date).
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- This dataset is a structured sample derived from larger repositories (e.g., national or regional health databases). While comprehensive, it may not include every instance globally.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance includes raw counts (e.g., cases, deaths), demographic variables (age, region, sex), and calculated fields like CFR. Data preprocessing includes cleaning for missing or invalid entries.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- While not directly labeled, "high risk" categories were derived during analysis based on the CFR and median thresholds.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Certain entries might have missing data for specific regions or demographics due to reporting inconsistencies. There were also some problematic datapoints where the number of deaths for a specific demographic exceeded the number of cases. 
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- TBD
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- TBD
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- TBD
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- TBD
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- TBD
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- TBD
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- TBD
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- TBD
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- TBD
16. *Any other comments?*
	- TBD

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- Data was collected from government health agencies, WHO reports, and academic repositories. Preprocessing ensured standardization.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Collection involved web scraping, manual curation, and automated aggregation pipelines (e.g., through APIs).
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset is a near-comprehensive aggregation but may have regional gaps due to reporting biases or lack of publicly available data.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Preprocessing included: Removal of invalid entries (e.g., deaths exceeding cases), creation of age intervals, calculation of metrics like CFR, and filtering for specific regions, sexes, and age groups.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References

### 2024-11-26 paper/paper.qmd
Updated paper, models, and citations

@@
 title: "Number of COVID-19 Cases in Canada By Region, Age, and Sex"
-subtitle: "Predicting the Factors that Influence the Number of COVID-19 Cases and Deaths"
+subtitle: "Predicting the Factors that Influence the Probability of High Risk of COVID-19"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper examines the demographic and regional factors influencing high-risk classifications for COVID-19 in Canada using logistic regression models, with predictors including age, sex, and region.
+The analysis reveals that rural regions, males, and specific age intervals exhibit higher probabilities of being classified as high risk, emphasizing disparities in healthcare access and outcomes.
+These findings highlight the importance of targeted public health interventions, such as resource allocation and vaccination prioritization, to mitigate these risks and can potentially guide interventions for future pandemics.
+By identifying key predictors and their implications, this study provides a scalable framework for improving pandemic preparedness and understanding health disparities."
 format: pdf
+toc: true
+toc-depth: 3
+include-before:
+  - \newpage
 number-sections: true
@@
 ```
-
+\newpage
 # Introduction
@@
 
-Results paragraph (what was found?)
+The analysis revealed that there are disparities in the risk of severe COVID-19 outcomes across demographic and regional categories.
+Logistic regression models demonstrated that individuals in rural regions consistently faced higher probabilities of being classified as high risk, with males and specific age intervals further amplifying this vulnerability.
+These findings were supported by consistent trends in summary statistics and model visualizations, which revealed the interplay between predictors such as age, region, and sex in influencing COVID-19 outcomes.
 
-Why it matters paragraph (why it matters?)
+Understanding the drivers of COVID-19 severity can craft equitable and effective public health responses.
+By identifying which demographic and regional groups face the greatest risks, the results of the analysis can guide resource distribution, including vaccination programs and medical support, to areas where they are most needed.
+These are applicable not only to COVID-19 but also to future public health challenges.
 
 The remainder of this paper is structured as follows.
-@sec-data highlights the characteristics of the dataset.
+The data section (@sec-data) highlights the characteristics of the dataset.
 The section summarizes the data through a series of summary statistics (@fig-summary) and represents the data in a visual way where specific trends and patterns can be observed (@fig-cases_over_time, @fig-cfr).
-@sec-model creates a logistic regression model predicting the probability of high risk depending on the region, sex, and age of the data point.
-The section includes model figures that display the predicted probabilities of high risk for the predictor variables respectively.
+The model section (@sec-model) creates a logistic regression model predicting the probability of high risk depending on the region, sex, and age of the data point.
+This section presents the mathematical basis of the model and justifies the selection for the logistic regression model over others.
+The results section (@sec-results) includes model figures that display the predicted probabilities of high risk for the predictor variables respectively.
+The discussion section (@sec-discussion) goes over how the results from the modeling can be interpreted in the context of the real world and how these findings can be utilized to advise public health measures.
+It also discusses limitations of the analysis and future directions.
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The dataset reflects a transformation of real-world phenomena—COVID-19 cases, deaths, and testing rates—into structured entries that can be analyzed.
+Each entry in the dataset represents aggregated information collected from health departments, governmental organizations, and research institutions worldwide.
+These organizations report COVID-19 statistics at varying levels of granularity, such as daily case counts or weekly summaries, which are then standardized and compiled into the dataset.
+This collective and comprehensive effort of data collection allows the real-world challenges of facing COVID-19 to be transformed into quantitative measures that can be analyzed and modeled.
 
@@
   ) +
+  scale_fill_manual(
+    values = c("m" = "steelblue", "f" = "pink") # Specify colors for "m" and "f"
+  ) +
   theme_minimal() +
@@
   geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with CFR
-  scale_fill_manual(values = c("Urban" = "steelblue", "Rural" = "forestgreen")) + # Optional: Region-specific colors
+  scale_fill_manual(values = c("Urban" = "lightgreen", "Rural" = "yellow")) + # Optional: Region-specific colors
   labs(
@@
 
-## Model set-up
+## Model set-up {#sec-model_setup}
 
@@
 
-The \textbf{box-and-whisker plot} visualizes the distribution of predicted probabilities (\( p \)) across different \textbf{age intervals}.
+The box-and-whisker plot visualizes the distribution of predicted probabilities (\( p \)) across different age intervals.
 Each box represents the interquartile range (IQR) of predicted probabilities, with the line inside the box indicating the median predicted probability for that age interval.
@@
 \begin{align}
-\text{logit}(p) = \ln\left(\frac{p}{1 - p}\right) = \beta_0 + \beta_1 \cdot \text{Tests} + \beta_2 \cdot \text{Age Interval} + \epsilon
+\text{logit}(p) = \ln\left(\frac{p}{1 - p}\right) = \beta_0 + \beta_1 \cdot \text{Tests} + \beta_2 \cdot \text{Age} + \beta_3 \cdot \text{Region} + \epsilon
 \end{align}
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We run the model in R [@citeR] using the 'pROC' [@pROC] and 'caret [@caret] packages to construct and test the model.
 
-### Model justification
+## Model justification {#sec-model_justification}
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+For the analysis, a logistic regression model was chosen to predict the probability of a region or individual being classified as "high risk" based on key predictors such as age, sex, region, and the number of COVID-19 tests conducted.
+Logistic regression is particularly well-suited for this dataset because the outcome variable, high risk, is binary (e.g., classified as high risk or not).
+This makes logistic regression the most appropriate modeling technique as it directly estimates the probability of the binary outcome through the log-odds transformation.
+Compared to other models like linear regression, logistic regression avoids the issue of nonsensical predictions, such as probabilities below 0 or above 1, ensuring the interpretability of the results.
+Most importantly, logistic regression can handle both categorical and continuous predictors effectively.
+For this particular dataset, the key predictor variables are all categorical, even age is categorized by intervals.
+Therefore, it was extremely necessary for the model to be able to handl categorical data, which the logistic regression model is capable of.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+Another key advantage of the logistic regression model in this context is that it provides interpretable coefficients, which allow for an understanding of how specific predictors influence the odds of being high risk.
+This interpretability is crucial for public health applications, where stakeholders need clear and actionable insights.
+Furthermore, logistic regression is computationally efficient, making it ideal for datasets with relatively straightforward structures like this one.
+Thus, the logistic regression model strikes an appropriate balance between accuracy, interpretability, and simplicity, making it the most suitable choice for this analysis.
+
+The details on model selection can be found in the appendix in @sec-appendixb
+
+## Assumptions and Limitations
+
+The logistic regression model assumes that observations within the dataset are independent, meaning that each data point does not influence others.
+However, this assumption could be violated if individuals within the same region share similar environmental, healthcare, or socioeconomic factors that contribute to their risk classification.
+Additionally, the model assumes that the predictors, such as age, sex, and region, are not highly correlated.
+Significant multicollinearity between these variables could undermine the stability of the model coefficients and reduce interpretability.
+Finally, the classification of "high risk" as regions or individuals exceeding the median death rate simplifies a complex phenomenon.
+This threshold may fail to capture subtler variations in risk factors across different demographic and geographic contexts.
+
+One key limitation of logistic regression is its inability to account for nuanced interactions between predictors without explicitly adding interaction terms.
+For instance, the combined effects of age and sex on risk may not be fully captured in this model.
+Additionally, the model assumes a predefined functional form, which limits its flexibility in identifying nonlinear relationships between predictors and outcomes.
+As a result, more complex patterns in the data may remain undetected, suggesting the need for alternative modeling approaches such as random forests or decision trees in future studies.
+
+## Model Validation
+
+For the model validation process of the selected logistic regression model, multiple validation techniques were used.
+Firstly, a ROC curve was constructed to visualize the trade-off between sensitivity and specificity.
+The ROC curve showed a high curve that deviated from the linear line representing random chance between sensitivity and specificity.
+The AUC value accompanying the ROC curve was above 0.8 for all three prediction models constructed for different variables.
+
+A confusion matrix was also constructed to assess classification accuracy, precision, and recall by comparing the predicted classes to the actual outcomes.
+Values of above 0.75 were yielded for the accuracy rating.
+Lastly, the odds ratios with confidence intervals provided insights into the strngth and direction of the predictors' effects.
+
+In general, these results demonstrated the suitability and strong effect that the logistic regression model has to predict probability of high risk based on the data.
+
+Details of the model validation are provided in the appendix (@sec-appendixc).
 
@@
 
-Our results are summarized in @tbl-modelresults.
+The results of this analysis are summarized in @fig-model_region, @fig-model_sex, and @fig-model_age to give an all-rounded picture of the predicted probability of high risk based on the different predictor variables.
+
+## Predicted Probability of High Risk by Region
 
 ```{r}
-#| label: model-risk_urban_vs_rural
+#| label: fig-model_region
 #| fig-cap: Predicted probability of high risk of COVID-19 by region (Urban vs. Rural)
@@
 #| message: false
+#| results: 'hide'
 
@@
 logistic_model <- glm(
-  High_Risk ~ Tests + Region + Age,
+  High_Risk ~ Sex + Region + Age + Cases,
   family = binomial(),
@@
 
-# Step 4: Visualize the Results
-# Plot predicted probabilities by region
-ggplot(filtered_data, aes(x = Region, y = Predicted_Probability, fill = Region)) +
-  geom_boxplot(alpha = 0.6) +
+# Visualization with Density Plots
+ggplot(filtered_data, aes(x = Predicted_Probability, fill = Region)) +
+  geom_density(alpha = 0.6) + # Density plot for predicted probabilities
+  facet_wrap(~Region, scales = "free") + # Separate density plots by region
   labs(
-    title = "",
-    x = "Region",
-    y = "Predicted Probability of High Risk"
+    title = "Density of Predicted Probability of High Risk by Region",
+    x = "Predicted Probability of High Risk",
+    y = "Density"
+  ) +
+  scale_fill_manual(
+    values = c("Urban" = "lightgreen", "Rural" = "yellow") 
   ) +
@@
 
+@fig-model_region models the predicted probability of high risk for different regions, urban or rural. This is represented by a faceted density plot.
+
+Based on these results, the predicted probability of high risk is higher in rural areas than urban areas and the predicted probability of low risk is higher in urban areas than rural.
+This is indicated by the height of the peaks at the higher and lower end of the scale, where the peaks for urban regions are taller at the lower end of the probability of high risk scale and shorter at the higher end.
+The height of the peaks indicate the density of datapoints that fall around the respective predicted value.
+
+\newpage
+## Predicted Probability of High Risk by Sex
+
 ```{r}
-#| label: model-risk_sex
+#| label: fig-model_sex
 #| fig-cap: Predicted probability of high risk based on sex
@@
 logistic_model <- glm(
-  High_Risk ~ Tests + Sex + Age,
+  High_Risk ~ Sex + Sex + Age + Cases,
   family = binomial(),
@@
   ) +
+  scale_fill_manual(
+    values = c("m" = "steelblue", "f" = "pink") # Specify colors for "m" and "f"
+  ) +
   theme_minimal()
@@
 
+@fig-model_sex compares the predicted probability of high risk based on biological sex, representing the model in a similar box-and-whiskers plot method.
+As shown in the figure, males (represented by 'm') have a noticeably higher predicted probability of high risk than females (represented by 'f').
+The predicted probability of a male qualifying as high risk for COVID-19 is above 50%, while the number is below 50% for females.
+
+## Predicted Probability of High Risk by Age Interval
+
 ```{r}
-#| label: model-risk_age
+#| label: fig-model_age
 #| fig-cap: Predicted probability of high risk based on age intervals
@@
 logistic_model <- glm(
-  High_Risk ~ Tests + Age_Interval,
+  High_Risk ~ Sex + Region + Age_Interval + Cases,
   family = binomial(),
@@
 
-# Step 4: Visualize the Results
-# Plot predicted probabilities by age interval
-ggplot(filtered_data, aes(x = Age_Interval, y = Predicted_Probability, fill = Age_Interval)) +
-  geom_boxplot(alpha = 0.6) +
+# Compute Median Predicted Probabilities
+median_data <- filtered_data %>%
+  group_by(Age_Interval) %>%
+  summarise(Median_Probability = median(Predicted_Probability, na.rm = TRUE))
+
+# Visualization with Median Bar Plot (Single Color and No Legend)
+ggplot(median_data, aes(x = Age_Interval, y = Median_Probability)) +
+  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) + # All bars have the same color
+  geom_text(aes(label = round(Median_Probability, 2)), vjust = -0.5) + # Add value labels
   labs(
@@
     x = "Age Interval",
-    y = "Predicted Probability of High Risk"
+    y = "Median Predicted Probability of High Risk"
   ) +
+  ylim(0, 0.65) +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for readability
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels for readability
+    legend.position = "none" # Remove the legend
+  )
 
@@
 
-# Discussion
+@fig-model_age shows the predicted probability of high risk for different age groups.
+The age groups are intervals of 10 from 0 to 100 as classified by the original dataset.
+As shown in the figure, age intervals 10 to 20 and 30 to 40 have a significantly higher predicted probability of high risk than other age groups.
+This means that it is more likely (around 55% and 57% chance) that someone within these age ranges will qualify as a high risk than the other age groups.
+The age group with the lowest predicted probability of high risk is 0 to 10 year-olds, yielding a 38% chance median of being predicted as high risk.
 
-## First discussion point {#sec-first-point}
+# Discussion {#sec-discussion}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+## Interpreting High Risk Predictions {#sec-firstpoint}
 
-## Second discussion point
+The results of this analysis reveal critical insights into how demographic and regional factors influence the likelihood of being classified as high risk for COVID-19.
+The logistic regression model, designed to predict the probability of high risk, effectively highlighted the relative importance of age, sex, and region in shaping these probabilities.
+Notably, rural regions consistently showed higher predicted probabilities of high risk compared to urban regions.
+This disparity underscores the challenges faced by rural areas, such as limited access to healthcare infrastructure, which may contribute to higher case fatality rates (CFR).
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+Similarly, the analysis revealed that males are more likely to be classified as high risk than females, aligning with previous research suggesting that biological differences, comorbidities, and health-seeking behaviors could account for this discrepancy.
+Age also played a pivotal role, with certain age intervals, such as 10–20 and 30–40, showing significantly higher probabilities of high risk.
+This pattern may reflect complex interactions between behavioral, immunological, and societal factors influencing COVID-19 outcomes.
+Understanding these relationships is crucial for tailoring public health interventions and resource allocation strategies.
 
-## Third discussion point
+## Key Future Implications {#sec-secondpoint}
 
-## Weaknesses and next steps
+The findings of this study have implications for future pandemic preparedness and response.
+By identifying demographic and regional factors associated with higher probabilities of high risk, policymakers can develop targeted interventions to mitigate disparities.
+For example, enhancing healthcare infrastructure in rural areas and prioritizing vaccination and testing campaigns for high-risk groups can help reduce the impact of future pandemics.
 
-Weaknesses and next steps should also be included.
+Additionally, the use of logistic regression models to predict high-risk classifications provides a robust framework for decision-making.
+This approach could be extended to other diseases or health outcomes, offering a scalable and interpretable method for public health planning.
+The analysis also highlights the importance of maintaining high-quality, standardized datasets to facilitate cross-regional and cross-demographic comparisons.
+
+## Limitations {#sec-thirdpoint}
+
+Despite its strengths, this study has several limitations that should be addressed in future research.
+First, the analysis is limited to Canada, which may restrict the generalizability of the findings to other countries with different healthcare systems, cultural contexts, or pandemic responses.
+Second, the dataset used for this analysis only includes data from 2022, which may not capture temporal variations in COVID-19 patterns.
+Future studies could incorporate longitudinal data to examine how the predictors of high risk evolve over time.
+
+Another limitation is the reliance on aggregate data, which may obscure individual-level variability and interactions.
+For example, the interplay between age, sex, and comorbidities at the individual level could yield more nuanced insights.
+Additionally, the classification of rural and urban regions may oversimplify the diversity within these categories, such as the distinction between suburban and remote rural areas.
+
+Thirdly, the data obtained for COVID-19 is obviously in the past and no longer an issue that is of primary concern.
+This issue is in line with the generalizability issue due to limited surveying countries.
+The patterns and models may be limited in generalizability due to the sole focus on COVID-19.
+
+## Next Steps {#sec-fourthpoint}
+
+Building on the findings of this study, future research should explore several directions.
+First, expanding the analysis to include additional countries or regions would provide a more comprehensive understanding of the factors influencing COVID-19 outcomes.
+Second, incorporating additional variables, such as socioeconomic status, vaccination rates, and comorbidities, could enhance the predictive power of the models.
+
+Another next step that could be taken is to use the analysis and modeling on a broader series of pandemics and diseases such as seasonal flues, SARS outbreak in the 2000s, and others that may show similar trends.
+This will allow a broader generalizability of the demographic patterns and allow the model to better predict probability of high risk based on demographic data.
+Finally, public health practitioners and researchers should collaborate to ensure that findings are translated into actionable policies and interventions, bridging the gap between data analysis and real-world impact.
 
@@
 
-# Appendix {.unnumbered}
+# Appendix {#sec-appendix}
 
-# Additional data details
+## Appendix A: Survey and Sampling Details {#sec-appendixa}
 
-# Model details {#sec-model-details}
+The dataset used in this analysis is sourced from the COVerAGE-DB, a comprehensive repository of COVID-19 demographic data.
+The database compiles data from a diverse set of sources, including government health agencies, international organizations, and other reputable institutions, ensuring a broad representation of regions and demographics.
+Each entry in the dataset represents an aggregation of cases, deaths, and demographic information categorized by age, sex, and region.
 
-## Posterior predictive check
+### Collection Process
+
+Data collection follows a systematic approach.
+For primary sources, COVerAGE-DB retrieves data directly from official government dashboards, public health reports, and institutional publications (@cite-dowd2020).
+This ensures that the dataset reflects real-world phenomena with minimal distortion.
+
+The database ensures standardization by processing the raw data from various sources through rigorous standardization to ensure comparability.
+For example, age intervals are harmonized across sources to create a unified structure suitable for cross-regional analysis.
+
+The database also ensures quality Control through running automated checks and manual reviews to detect inconsistencies, missing values, and potential errors in data entry.
+Such measures are crucial given the heterogeneity in reporting standards across countries and regions.
+
+Lastly, the update frequency is ensured so the data is as relevant and recent as possible.
+The database is updated regularly to reflect new data releases, ensuring timeliness and relevance for ongoing research.
+
+### Sampling and Representativeness
+
+While the dataset covers a broad range of countries and regions, its representativeness depends on the completeness and accuracy of the underlying sources.
+For instance, regions with underdeveloped health reporting infrastructure might have lower data quality or incomplete records.
+Additionally, variations in testing capacity and reporting standards can introduce biases, particularly in the calculation of derived measures such as case fatality rates (CFR).
+
+### Observational Nature and Limitations
+
+The data is observational, collected without experimental controls.
+As a result, confounding factors such as healthcare system differences, socioeconomic disparities, and public health policies must be carefully considered when interpreting results (@cite-ioannidis2021).
+For instance, differences in testing rates across regions can influence the apparent number of cases and consequently impact measures like CFR.
+
+## Appendix B: Model Selection Details {#sec-appendixb}
+
+During the model selection process, model figures were generated for the linear regression model.
+As shown in @fig-appendixb, the data was difficult to represent and relationships were difficult to identify due to the categorical nature of the variables.
+This led to attempts to model the data in different ways.
+The second method that was attempted was using a logistic regression model, which was able to represent the data well and make predictions about high risk.
+These figures are shown within the @sec-results of the paper.
 
 ```{r}
+#| label: fig-appendixb
+#| fig-cap: Linear Regression Model for dataset
+#| echo: false
+#| warning: false
+#| message: false
+#| results: 'hide'
+
+# Load necessary libraries
+library(dplyr)
+library(ggplot2)
+set.seed(21)
+
+# Prepare the data by calculating CFR and filtering for valid rows
+filtered_data <- analysis_data %>%
+  mutate(CFR = (Deaths / Cases) * 100) %>% # Calculate CFR (%)
+  filter(!is.na(CFR)) # Remove rows with missing CFR
+
+# Fit a linear regression model
+linear_model <- lm(CFR ~ Age + Sex + Region, data = filtered_data)
+
+# Summary of the linear regression model
+summary(linear_model)
+
+# Visualize the model predictions vs actual CFR
+filtered_data <- filtered_data %>%
+  mutate(Predicted_CFR = predict(linear_model, newdata = filtered_data))
+
+ggplot(filtered_data, aes(x = Predicted_CFR, y = CFR)) +
+  geom_point(alpha = 0.5) +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
+  labs(
+    title = "Predicted vs Actual CFR (Linear Regression)",
+    x = "Predicted CFR",
+    y = "Actual CFR"
+  ) +
+  theme_minimal()
 
@@
 
-## Diagnostics
+## Appendix C: Model Validation Details {#sec-appendixc}
+
+### ROC Curve and AUC
 
 ```{r}
+#| label: fig-appendix_roc
+#| fig-cap: ROC Curve and AUC Value for Logistic Regression Model
+#| echo: false
+#| warning: false
+#| message: false
 
+# Load necessary libraries
+library(pROC)
+library(ggplot2)
+set.seed(21)
+
+analysis_data <- read.csv(here("data/02-analysis_data/modeling_data.csv"))
+
+# Ensure High_Risk is binary (convert if necessary)
+analysis_data <- analysis_data %>%
+  mutate(High_Risk = as.numeric(as.character(High_Risk)))
+
+# Generate the ROC curve
+roc_curve <- roc(
+  response = analysis_data$High_Risk,  # True labels
+  predictor = analysis_data$Predicted_Probability  # Predicted probabilities
+)
+
+# Calculate AUC
+auc_value <- auc(roc_curve)
+
+# Plot the ROC curve
+roc_plot <- ggplot() +
+  geom_line(data = data.frame(
+    specificity = rev(roc_curve$specificities),
+    sensitivity = rev(roc_curve$sensitivities)
+  ), aes(x = specificity, y = sensitivity), color = "blue", size = 1) +
+  labs(
+    title = "ROC Curve",
+    x = "Specificity",
+    y = "Sensitivity"
+  ) +
+  theme_minimal()
+
+# Display the AUC value
+cat("AUC:", auc_value, "\n")
+
+# Generate the ROC curve
+roc_curve <- roc(analysis_data$High_Risk, analysis_data$Predicted_Probability)
+auc_value <- auc(roc_curve)
+
+# Plot the ROC curve
+roc_plot <- ggplot() +
+  geom_line(
+    data = data.frame(
+      specificity = rev(roc_curve$specificities),
+      sensitivity = rev(roc_curve$sensitivities)
+    ),
+    aes(x = specificity, y = sensitivity),
+    color = "black", size = 1.2
+  ) +
+  geom_abline(linetype = "dashed", color = "red") + # Add the diagonal line
+  labs(
+    title = "",
+    x = "1 - Specificity",
+    y = "Sensitivity"
+  ) +
+  annotate("text", x = 0.7, y = 0.2, label = paste0("AUC = ", round(auc_value, 3)), size = 5, color = "black") +
+  theme_minimal() +
+  theme(
+    axis.title = element_text(size = 14),
+    plot.title = element_text(size = 16, face = "bold")
+  )
+
+# Print the plot in your script to display in the rendered document
+roc_plot
+
+```
+
+### Confusion Matrix
+
+```{r}
+#| label: fig-appendix_confusion
+#| fig-cap: Confusion Matrix for Logistic Regression Model
+#| echo: false
+#| warning: false
+#| message: false
+
+# Load necessary library
+library(caret)
+set.seed(21)
+
+# Step 1: Create Predicted Classes
+# Add a column with predicted classes based on a 0.5 threshold
+analysis_data <- analysis_data %>%
+  mutate(
+    Predicted_Class = ifelse(Predicted_Probability >= 0.5, 1, 0) # Classify as High Risk (1) if probability >= 0.5
+  )
+
+# Step 2: Generate Confusion Matrix
+# Convert classes to factors for compatibility
+actual_classes <- factor(analysis_data$High_Risk, levels = c(0, 1))
+predicted_classes <- factor(analysis_data$Predicted_Class, levels = c(0, 1))
+
+# Create confusion matrix
+conf_matrix <- confusionMatrix(predicted_classes, actual_classes)
+
+# Step 3: Print Confusion Matrix
+print(conf_matrix)
+
+```
+
+### Odds Ratio with Confidence Intervals
+
+```{r}
+#| label: fig-odds_ratio
+#| fig-cap: Odds Ratio table with Confidence Intervals
+#| echo: false
+#| warning: false
+#| message: false
+
+# Load necessary libraries
+library(broom)
+library(knitr)
+library(dplyr)
+set.seed(21)
+
+# Step 1: Compute Odds Ratios and Confidence Intervals
+# Use `tidy()` to extract model coefficients and calculate odds ratios
+odds_ratios <- tidy(logistic_model, conf.int = TRUE, exponentiate = TRUE)
+
+# Step 2: Format the Table for Better Display
+# Rename columns for better readability
+odds_ratios <- odds_ratios %>%
+  rename(
+    Predictor = term,
+    Odds_Ratio = estimate,
+    Lower_95CI = conf.low,
+    Upper_95CI = conf.high
+  )
+
+# Step 3: Display the Table using kable
+kable(
+  odds_ratios,
+  col.names = c("Predictor", "Odds Ratio", "Standard Error", "Statistic", "P Value", "Lower 95% CI", "Upper G5% CI"), 
+  booktabs = TRUE,
+  digits = 3
+)
 ```

### 2024-11-26 scripts/06-model_data.R

@@ -16,10 +16,19 @@ library(dplyr)
 library(caret)
 library(pROC)
 
+
 # Building a logistic regression model for predicted high risk and age
+# Step 0: Define High_Risk variable in the dataset
+analysis_data <- analysis_data %>%
+  mutate(
+    Death_Rate = Deaths / (Cases + 1e-6), # Calculate death rate (avoid division by zero)
+    Log_CFR = log(CFR + 1),
+    High_Risk = ifelse(Death_Rate > median(Death_Rate, na.rm = TRUE), 1, 0) # Classify high-risk regions
+  )
+
 # Step 1: Build the logistic regression model
 logistic_model <- glm(
-  High_Risk ~ Age + Region + Sex,  # Replace with relevant predictors
+  High_Risk ~ Age + Region + Sex + Cases, # Replace with relevant predictors
   data = analysis_data,
   family = binomial()
 )
@@ -101,7 +110,7 @@ filtered_data <- analysis_data %>%
 
 # Step 2: Fit Logistic Regression Model
 logistic_model <- glm(
-  High_Risk ~ Tests + Sex + Age,  # Predictors: Tests, Sex, Age
+  High_Risk ~ Tests + Sex + Age + Cases,  # Predictors: Tests, Sex, Age
   family = binomial(),
   data = filtered_data
 )
@@ -159,7 +168,7 @@ filtered_data <- analysis_data %>%
 
 # Step 2: Fit Logistic Regression Model
 logistic_model <- glm(
-  High_Risk ~ Tests + Region + Age,  # Predictors: Tests, Region, Age
+  High_Risk ~ Tests + Region + Age + Cases,  # Predictors: Tests, Region, Age
   family = binomial(),
   data = filtered_data
 )
@@ -204,3 +213,6 @@ ggplot(filtered_data, aes(x = Region, y = Predicted_Probability, fill = Region))
 
 # Step 6: Save the Model for Reproducibility
 saveRDS(logistic_model, file = "models/logistic_model_region_high_risk.rds")
+
+write.csv(analysis_data, "data/02-analysis_data/modeling_data.csv", row.names = FALSE)
+

### 2024-11-26 paper/paper.qmd
Updated prose of paper

@@
 The COVID-19 pandemic is something that is difficult to forget.
-Although it is now in the past, the effects that it had on society and how we function as individuals is still profound.
+Although it is now in the past, the effects that it had on society and how we function as individuals is still noticeable.
 At the same time, the statistical information that sprouted from the pandemic has a high value for analysis.
@@
 There has also not been cross comparisons of region and demographic aspects.
-Another gap is the lack of country-specific analysis, particularly Canada.
+Another area where research is needed is for country-specific analysis, particularly Canada.
 
@@
 
-The analysis revealed that there are disparities in the risk of severe COVID-19 outcomes across demographic and regional categories.
+The analysis showed that there are disparities in the risk of severe COVID-19 outcomes across demographic and regional categories.
 Logistic regression models demonstrated that individuals in rural regions consistently faced higher probabilities of being classified as high risk, with males and specific age intervals further amplifying this vulnerability.
-These findings were supported by consistent trends in summary statistics and model visualizations, which revealed the interplay between predictors such as age, region, and sex in influencing COVID-19 outcomes.
+These findings were supported by consistent trends in summary statistics and model visualizations, which showed the interplay between predictors such as age, region, and sex in influencing COVID-19 outcomes.
 
-Understanding the drivers of COVID-19 severity can craft equitable and effective public health responses.
+Understanding the reasons behind COVID-19 severity can craft equitable and effective public health responses.
 By identifying which demographic and regional groups face the greatest risks, the results of the analysis can guide resource distribution, including vaccination programs and medical support, to areas where they are most needed.
@@
 These organizations report COVID-19 statistics at varying levels of granularity, such as daily case counts or weekly summaries, which are then standardized and compiled into the dataset.
-This collective and comprehensive effort of data collection allows the real-world challenges of facing COVID-19 to be transformed into quantitative measures that can be analyzed and modeled.
+This effort of data collection allows the real-world challenges of facing COVID-19 to be transformed into quantitative measures that can be analyzed and modeled.
 
@@
 Another key advantage of the logistic regression model in this context is that it provides interpretable coefficients, which allow for an understanding of how specific predictors influence the odds of being high risk.
-This interpretability is crucial for public health applications, where stakeholders need clear and actionable insights.
+This interpretability is crucial for public health applications.
 Furthermore, logistic regression is computationally efficient, making it ideal for datasets with relatively straightforward structures like this one.
@@
 
-One key limitation of logistic regression is its inability to account for nuanced interactions between predictors without explicitly adding interaction terms.
+One key limitation of logistic regression is its inability to account for interactions between predictors without explicitly adding interaction terms.
 For instance, the combined effects of age and sex on risk may not be fully captured in this model.
@@
 Values of above 0.75 were yielded for the accuracy rating.
-Lastly, the odds ratios with confidence intervals provided insights into the strngth and direction of the predictors' effects.
+Lastly, the odds ratios with confidence intervals provided information on the strength and direction of the predictors' effects.
 
@@
 
-The results of this analysis reveal critical insights into how demographic and regional factors influence the likelihood of being classified as high risk for COVID-19.
+The results of this analysis displays information into how demographic and regional factors influence the likelihood of being classified as high risk for COVID-19.
 The logistic regression model, designed to predict the probability of high risk, effectively highlighted the relative importance of age, sex, and region in shaping these probabilities.
@@
 This pattern may reflect complex interactions between behavioral, immunological, and societal factors influencing COVID-19 outcomes.
-Understanding these relationships is crucial for tailoring public health interventions and resource allocation strategies.
+Understanding these relationships is important for tailoring public health interventions and resource allocation strategies.
 
@@
 Building on the findings of this study, future research should explore several directions.
-First, expanding the analysis to include additional countries or regions would provide a more comprehensive understanding of the factors influencing COVID-19 outcomes.
+First, expanding the analysis to include additional countries or regions would provide a better understanding of the factors influencing COVID-19 outcomes.
 Second, incorporating additional variables, such as socioeconomic status, vaccination rates, and comorbidities, could enhance the predictive power of the models.
@@
 This will allow a broader generalizability of the demographic patterns and allow the model to better predict probability of high risk based on demographic data.
-Finally, public health practitioners and researchers should collaborate to ensure that findings are translated into actionable policies and interventions, bridging the gap between data analysis and real-world impact.
+Finally, public health practitioners and researchers should collaborate to ensure that findings are translated into actionable policies and interventions.
 
@@
 
-The dataset used in this analysis is sourced from the COVerAGE-DB, a comprehensive repository of COVID-19 demographic data.
+The dataset used in this analysis is sourced from the COVerAGE-DB, a repository of COVID-19 demographic data.
 The database compiles data from a diverse set of sources, including government health agencies, international organizations, and other reputable institutions, ensuring a broad representation of regions and demographics.

### 2024-11-26 other/datasheet/datasheet_covid_data.qmd
Updated sketches and data sheet

@@
 thanks: "Code and data are available at: LINK."
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+date: today
+date-format: long
+abstract: ""
 format: pdf
 number-sections: true
-bibliography: references.bib
+bibliography: references_data.bib
 ---
@@
 4. *Any other comments?*
-    - TBD
+    - The overall motivation of the data collection was pushed by the large social and political say that COVID-19 had and the importance to understand the pandemic overall.
 
@@
 7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
-	- TBD
+	- Instances are grouped by age, sex, and region but do not explicitly represent relationships between different demographic groups.
+The specific indicators of demographics are not necessarily correlated with each other.
 8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
-	- TBD
+	- The dataset can be split temporally (e.g., pre- and post-vaccine availability), geographically (e.g., urban vs.
+rural, e.g.
+by country), or by other demographic characteristics to test specific hypotheses.
+These can lead to different trends and patterns based on what the focus is for the analysis.
 9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
-	- TBD
+	- The dataset may include errors due to underreporting, misclassification, or inconsistencies in data collection. An example is datapoints having more number of deaths than number of cases. 
 10.
@@
 Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
-	- TBD
+	- The dataset relies on external reporting sources such as health departments and international organizations like WHO, with potential limitations in long-term availability.
 11.
@@
 If so, please provide a description.*
-	- TBD
+	- The dataset contains aggregated data and does not identify individuals, minimizing confidentiality concerns.
 12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
-	- TBD
+	- While the dataset itself is not offensive, findings could raise concerns about inequities or vulnerabilities in specific populations.
 13.
@@
 If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
-	- TBD
+	- Sub-populations are identified by country, age group, sex, and region.
 14.
@@
 If so, please describe how.*
-	- TBD
+	- The dataset does not allow for individual identification as it is aggregated.
 15.
@@
 If so, please provide a description.*
-	- TBD
+	- The dataset involves public health data, which can be sensitive but is anonymized and aggregated.
 16. *Any other comments?*
-	- TBD
+	- The data is collected from government and healthcare sources anonymously. The source of the data is generally trustworthy and does not violate privacy or confidenciality.
 
@@
 4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
-	- TBD
+	- Data collection involved government officials, researchers, and volunteers in health agencies.
 5.
@@
 If not, please describe the timeframe in which the data associated with the instances was created.*
-	- TBD
+	- Data spans the COVID-19 pandemic timeline, with regular updates. Specifically, the data covers the year of 2022. 
 6.
@@
 If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+	- The dataset used in this analysis was aggregated from publicly available data provided by health organizations, governments, and international agencies (e.g., WHO).
+As the data is aggregated and anonymized, ethical reviews specific to this dataset were not conducted by the authors of this analysis.
+However, ethical guidelines from the original sources, such as health agencies, likely applied during the data collection and reporting process.
 7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	- TBD
+	- The data was obtained from third-party sources, including official health department reports, COVID-19 dashboards, and repositories such as the COVerAGE-DB and WHO's publicly available data.
 8.
@@
 If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
-	- TBD
+	- As the data is aggregated and anonymized, individual notification was not applicable.
+Data collection practices adhered to reporting standards established by the respective health authorities or governments.
 9.
@@
 If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
-	- TBD
+	- The dataset does not involve individual-level data but aggregated public health statistics. Consent was not required since the data does not contain personally identifiable information (PII).
 10.
@@
 If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
-	- TBD
+	- Not applicable. The dataset is anonymized and aggregated, and does not involve individual consent.
 11.
@@
 If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+	- Given that the dataset is anonymized and aggregated, a formal data protection impact analysis specific to this analysis was not conducted.
+However, the original data providers (e.g., WHO, government health organizations) are likely to have conducted such assessments to ensure compliance with data protection laws and ethical standards.
 12. *Any other comments?*
-	- TBD
+	- The data accessed was through COVerAGE-DB, which is a third-party that collects data from government and medical institutes on COVID-19.
+The details on the sources and methods of collection of the data were limited.
 
@@
 If not, you may skip the remaining questions in this section.*
-	- Preprocessing included: Removal of invalid entries (e.g., deaths exceeding cases), creation of age intervals, calculation of metrics like CFR, and filtering for specific regions, sexes, and age groups.
+	- Preprocessing using R [@citeR] included: Removal of invalid entries (e.g., deaths exceeding cases), creation of age intervals, calculation of metrics like CFR, and filtering for specific regions, sexes, and age groups.
 2.
@@
 If so, please provide a link or other access point to the "raw" data.*
-	- TBD
+	- Yes. The raw data can be found at https://osf.io/43ucn 
 3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
-	- TBD
+	- Yes. The software used was R programming, available at https://www.R-project.org/ 
 4. *Any other comments?*
-	- TBD
+	- No
 
@@
 1. *Has the dataset been used for any tasks already? If so, please provide a description.*
-	- TBD
+	- Yes, the dataset has been used for multiple tasks, including the analysis of COVID-19 trends by demographic factors such as age, sex, and region.
+Researchers have employed it to assess the impact of COVID-19 across different populations and inform policy-making.
+For example, studies have used this dataset to identify high-risk groups and evaluate disparities in healthcare outcomes.
+The dataset has also been cited in repositories such as GitHub for further analytical work and data visualizations.
 2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
-	- TBD
+	- The COVerAGE-DB has a GitHub repository that provides access to the data and associated analyses: https://github.com/timriffe/covid_age [@Riffe2021covidage].
+The repository contains scripts for data processing and examples of research using the dataset.
+Additional references to papers and systems utilizing the dataset may be linked within this repository or cited in its documentation.
 3. *What (other) tasks could the dataset be used for?*
-	- TBD
+	- The dataset could support analyses of vaccine impact, regional healthcare disparities, and demographic-specific risk factors.
 4.
@@
 Is there anything a dataset consumer could do to mitigate these risks or harms?*
-	- TBD
+	- Yes, users should be aware of potential reporting biases and inconsistencies in the dataset.
+As the dataset relies on publicly available COVID-19 data from various countries, differences in testing rates, reporting methods, and definitions of cases and deaths may introduce variability.
+Additionally, certain regions may have incomplete or missing data, potentially impacting the reliability of demographic analyses.
+To mitigate risks, users should consider cross-referencing data with other sources and explicitly noting any limitations due to reporting discrepancies.
 5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
-	- TBD
+	- The dataset’s accuracy is contingent on reporting consistency and may not reflect real-time dynamics.
 6. *Any other comments?*
-	- TBD
+	- The dataset is a valuable resource for understanding the demographic impacts of COVID-19, but its use is constrained by the dynamic nature of the pandemic and the evolving accuracy of data reporting.
+Users should approach the dataset with a clear understanding of its limitations, ensuring that any conclusions drawn are contextualized appropriately within these constraints.
 
@@
 If so, please provide a description.*
-	- TBD
+	- Yes, the dataset is made publicly available through platforms like the Open Science Framework (OSF) [@covid_data] and GitHub.
+These platforms enable third parties, including researchers, policymakers, and healthcare organizations, to access the dataset.
 2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
-	- TBD
+	- The dataset is distributed through GitHub as downloadable files and through OSF. The COVerAGE-DB repository does not currently have a DOI, but OSF assigns unique identifiers to its hosted files.
 3. *When will the dataset be distributed?*
-	- TBD
+	- The dataset has already been distributed and is updated periodically. Users can access the latest versions on GitHub and OSF.
 4.
@@
 If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
-	- TBD
+	- The dataset is distributed under open-access terms, and users are expected to cite the COVerAGE-DB project when utilizing the data.
+Licensing terms are not explicitly mentioned on GitHub or OSF but generally follow the principles of open-source data sharing for academic and research purposes.
 5.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
-	- TBD
+	- No explicit IP-based restrictions are mentioned. The data is freely available for academic and research use, and there are no fees or additional restrictions imposed.
 6.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
-	- TBD
+	- No export controls or regulatory restrictions apply to the dataset. It is shared openly across international platforms for use by researchers globally.
 7. *Any other comments?*
-	- TBD
+	- No
 
@@
 1. *Who will be supporting/hosting/maintaining the dataset?*
-	- TBD
+	- Maintenance is typically overseen by the organization providing the dataset, so COVerAGE-DB and Open Science Framework (OSF) maintains the dataset. 
 2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
-	- TBD
+	- This information is not provided
 3. *Is there an erratum? If so, please provide a link or other access point.*
-	- TBD
+	- Currently, no official erratum is mentioned. However, updates or corrections to the dataset are communicated via the GitHub repository or the Open Science Framework (OSF).
 4.
@@
 If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
-	- TBD
+	- Datasets are updated periodically to reflect new data or correct errors. It was last updated on November 29th of 2023. The intervals of update and the contributors are not mentioned. 
 5.
@@
 If so, please describe these limits and explain how they will be enforced.*
-	- TBD
+	- The dataset anonymizes data and does not contain personally identifiable information.
+Retention policies depend on the original data sources, and contributors are expected to comply with privacy and data protection laws of their respective countries.
 6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
-	- TBD
+	- Older versions of the dataset are hosted on the Open Science Framework (OSF) and GitHub.
+New updates and changes are documented in release notes or change logs, ensuring transparency.
+Obsolescence, if it occurs, would likely be communicated through these platforms.
 7.
@@
 If so, please provide a description.*
-	- TBD
+	- Yes, contributors can collaborate via the GitHub repository.
+Pull requests and issues can be submitted for dataset augmentation or suggestions.
+Contributions are reviewed by the dataset maintainers to ensure consistency and accuracy.
+The GitHub repository also provides guidelines for contribution.
 8. *Any other comments?*
-	- TBD
+	- The dataset is a collaborative effort relying on contributions from multiple global sources.
+While it serves as a comprehensive resource for analyzing COVID-19 trends, its utility depends on the accuracy of the underlying data sources.
+Users are encouraged to exercise caution when making causal inferences and consider cross-referencing data with other sources.
 

### 2024-11-26 paper/paper.qmd
Updated references

@@
   labs(
-    title = "Density of Predicted Probability of High Risk by Region",
+    title = "",
     x = "Predicted Probability of High Risk",

### 2024-12-03 paper/paper.qmd
Updated discussion of paper and scripts

@@
 ---
-title: "Number of COVID-19 Cases in Canada By Region, Age, and Sex"
+title: "Number of COVID-19 Cases in Canada in 2022 By Region, Age, and Sex"
 subtitle: "Predicting the Factors that Influence the Probability of High Risk of COVID-19"
@@
 library(here)
+library(arrow)
 
 # Load the dataset
-analysis_data <- read.csv(here("data/02-analysis_data/analysis_data.csv"))
+analysis_data <- read_parquet(here("data", "02-analysis_data", "analysis_data.parquet"))
 ```
@@
 
-This analysis looks at COVID-19 data from 2022 in Canada and analyzes how the different population and individual-level variables, such as age, sex, and region of life, influenced the risk of being diagnosed with COVID and the risk of fatality.
+This analysis looks at COVID-19 data from 2022 in Canada and analyzes how the different population and individual-level demographic variables, such as age, sex, and region of life, influenced the risk of fatality from COVID-19.
 The paper uses the demographic data gathered from 2022 to predict the characteristics of populations and individuals that may be at higher risk of fatality once they have been diagnosed with COVID.
-This analysis may have broader implications for preventative measures for flues, other diseases, and pandemics in the future.
 
+One importance of this information to be present is mainly for Canada-specific situations and patterns that may not be seen in a global scale analysis.
+The geographic location, temperature, population distribution, ethnic diversity, and other features of Canada is incredibly unique.
+The patterns present in other countries cannot be generalized to a smaller scale.
+From a public health standpoint, global health data is key in understanding overarching trends of epidemiology, but country specific (sometimes community specific) analyses are key for creating predictions based on a suitably-sized dataset [@frumkin].
+
+Another key importance of this analysis is its focus on basic demographic factors rather than specific health factors.
+It is often that a health issue, COVID-19, is examined through the lens of other health factors [@swanson_verdugo].
+For example, how does the presence of another long-term disease influence the risk of COVID-19.
+Looking at surface level demographic information can provide information that is useful to a broader audience.
+Specific diseases may not be possessed by everyone, but demographic information is.
+This information is also known by individuals, which will help them identify the category they fall into and examine their risk of COVID-19 based on their demographic information.
+
+The estimand of this analysis is the correlation between demographic information, including age, sex, and region, and the predicted probability of high risk for COVID-19.
 The analysis focuses on the outcome variable of predicted probability of high risk, which is a function of number of cases and number of deaths per datapoint and is further defined in @sec-model.
@@
 
+This analysis may have broader implications for preventative measures for flues, other diseases, and pandemics in the future.
 Understanding the reasons behind COVID-19 severity can craft equitable and effective public health responses.
@@
 The data section (@sec-data) highlights the characteristics of the dataset.
-The section summarizes the data through a series of summary statistics (@fig-summary) and represents the data in a visual way where specific trends and patterns can be observed (@fig-cases_over_time, @fig-cfr).
+The section summarizes the data through a series of summary statistics and represents the data in a visual way where specific trends and patterns can be observed.
 The model section (@sec-model) creates a logistic regression model predicting the probability of high risk depending on the region, sex, and age of the data point.
@@
 I use the statistical programming language R [@citeR] to analyze the data and to create graphs and models.
-Th packages that were used include tidyverse [@tidyverse], tidyr [@tidyr], dplyr [@dplyr], caret [@caret], and pROC [@pROC].
+Th packages that were used include tidyverse [@tidyverse], tidyr [@tidyr], dplyr [@dplyr], arrow [@arrow], DataExplorer [@dataexplorer], corrplot [@corrplot], caret [@caret], and pROC [@pROC].
 My data [@covid_data] was obtained from COVerAGE-DB, which was housed in Open Science Framework.
@@
 
-The dataset used in this analysis was gathered by COVerAGE-DB, a database that focuses on COVID related data.
+The data collection process is conducted by hospitals and governmental institutions that get first-handed numbers of COVID-19 patient information.
+Individuals who have symptoms and choose to visit a hospital will get tested, this number will be reported as the number of tests.
+Those that test positive for COVID-19 in the hospitals will be reported as the number of cases.
+Those who report deaths to the government level and indicate that the cause is due to COVID-19 will be recorded as the number of deaths due to COVID.
+
+The data that was collected through hopistals and government institutions was gathered by COVerAGE-DB, a database that focuses on COVID related data.
+COVerAGE-DB consolidates this information into the dataset that was used in the analysis.
 COVerAGE-DB often gathers data relating to COVID such as the number of cases, deaths, tests, and vaccinations through governmental institutes such as health ministries and statistical offices.
@@
 
+The data collected through hospitals or governmental institutes only reflect individuals who have visited hospitals for their symptoms and tested positive for COVID-19 there.
+This poses a key limitation of potential inaccuracy of numbers due to the large portion of individuals who self-test PCR, visit PCR testing stations, or don't get tested and stay home despite having symptoms.
+These situations were common given the situation in Canada in 2022, where individuals were encouraged to stay home if symptoms were experienced rather than visit hospitals to report their results.
+
 The dataset reflects a transformation of real-world phenomena—COVID-19 cases, deaths, and testing rates—into structured entries that can be analyzed.
+It also reflects the data collection challenges and biases of the real world, potentially understating values.
 Each entry in the dataset represents aggregated information collected from health departments, governmental organizations, and research institutions worldwide.
 These organizations report COVID-19 statistics at varying levels of granularity, such as daily case counts or weekly summaries, which are then standardized and compiled into the dataset.
-This effort of data collection allows the real-world challenges of facing COVID-19 to be transformed into quantitative measures that can be analyzed and modeled.
+This effort of data collection allows the real-world challenges, as accurately as possible despite obvious limitations, of facing COVID-19 to be transformed into quantitative measures that can be analyzed and modeled.
 
@@
 For the purpose to focus on data in Canada for this analysis, the datapoints for other countries were removed.
-Further cleaning was done to remove any problematic datapoints such as when the number of deaths exceeded the number of cases.
+This was the initial step in data cleaning.
+
+Further data cleaning was undertaken to ensure the quality and integrity of the dataset.
+First, missing values were identified and handled by checking each column.
+Any missing values were flagged and removed.
+Second, the dataset was checked for duplicates to avoid redundant entries, ensuring each row represented a unique combination of key columns such as Country, Region, Date, and Age.
+Additionally, logical consistency checks were conducted to identify and correct any issues, such as negative values in Cases, Deaths, or Tests, or where Deaths exceeded Cases, which are not plausible in real-world data.
+Inconsistent or invalid values in categorical variables, such as Region or Sex, were also detected and addressed to match predefined valid categories.
+Finally, summary statistics and visualizations were generated to examine the distribution of numerical variables and detect any anomalies that could indicate data entry errors.
 
@@
 The cleaned dataset contains variables including the country, region, date, sex, and age where the number of cases, deaths, and tests were reported.
-@fig-head_analysis_data shows a sample of the dataset, presenting the first six rows of the analysis data.
+@sec-appendixraw shows a sample of the dataset, presenting the first six rows of the analysis data.
 The outcome variables highlighted include the total number of cases and the total number of deaths.
@@
 ```{r}
-#| label: fig-head_analysis_data
-#| fig-cap: First six rows of the dataset after cleaning
-#| echo: false
-#| warning: false
-#| message: false
-
-# Load the necessary library
-library(knitr)
-
-# Display the first few rows of analysis_data
-head_table <- head(analysis_data)
-
-# Render the table with kable
-kable(
-  head_table,
-  booktabs = TRUE,  # Add better table styling
-)
-```
-
-```{r}
 #| label: fig-summary
-#| fig-cap: Summary Statistics of COVID-19 Cases in Canada
+#| fig-cap: Summary Statistics of COVID-19 Cases in Canada.
+The table shows the averages and totals of key numerical variables including age, cases, tests, and deaths.
+It also calculates the case fatality rate (CFR) as number of deaths divided by number of cases.
 #| echo: false
@@
 #| label: fig-cases_over_time
-#| fig-cap: "Number of cases per month in 2022 in Canada"
+#| fig-cap: Number of COVID-19 cases per month in 2022 in Canada. The red trend line highlights the pattern of COVID-19 cases by month in Canada throughout 2022. 
 #| warning: false
@@
 #| label: fig-cfr
-#| fig-cap: Case Fatality Rate (CFR) by Age Group, Region, and Sex
+#| fig-cap: Case Fatality Rate (CFR) by Age Group, Region, and Sex.
+Urban and rural regions are separated into two subfigures, each characterizing male and females by color and displaying differences by age intervals.
 #| echo: false
@@
 #| label: fig-cfr_by_region
-#| fig-cap: "Case Fatality Rate (CFR) by Region in 2022 in Canada"
+#| fig-cap: "Case Fatality Rate (CFR) by Region in 2022 in Canada. CFR percentage is shown to be slightly higher in urban areas than rural areas."
 #| warning: false
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+The primary goal of the model is to act as a tool for prediction for the level of high risk based on the characteristics of the specific datapoint.
+It can be used to identify how at risk a specific region, sex, or age group is for COVID-19 fatality.
+The mathematical definition of high risk is included in @sec-model_setup alongside a mathematical definition of CFR.
+Overall, the CFR is the number of deaths per cases while the high risk is when that level exceeds the median across the entire population.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The model described here is a logistic regression model.
+This was selected based on the binary and categorical nature of many predictor and outcome variables.
+Justification of this choice is included in @sec-model_justification and @sec-model_validation, with more details included in the appendix (@sec-appendixc).
 
@@
 
-## Assumptions and Limitations
+## Assumptions and Limitations {#sec-model_assumptions}
 
@@
 
-## Model Validation
+## Model Validation {#sec-model_validation}
 
@@
 
-## Predicted Probability of High Risk by Region
+## Predicted Probability of High Risk by Region {#sec-results_region}
 
@@
 #| label: fig-model_region
-#| fig-cap: Predicted probability of high risk of COVID-19 by region (Urban vs. Rural)
+#| fig-cap: Predicted probability of high risk of COVID-19 by region (Urban vs.
+Rural).
+This is a density plot showing that the extremes of predicted probability of high risk are more frequent for both urban and rural regions.
 #| echo: false
@@
 \newpage
-## Predicted Probability of High Risk by Sex
+## Predicted Probability of High Risk by Sex {#sec-results_sex}
 
@@
 #| label: fig-model_sex
-#| fig-cap: Predicted probability of high risk based on sex
+#| fig-cap: Predicted probability of high risk based on sex.
+The box-and-whiskers plots show that the median value of the predicted probability of high risk is similar for males and females.
+There is a large IQR for both.
 #| echo: false
@@
 
-## Predicted Probability of High Risk by Age Interval
+## Predicted Probability of High Risk by Age Interval {#sec-results_age}
 
@@
 #| label: fig-model_age
-#| fig-cap: Predicted probability of high risk based on age intervals
+#| fig-cap: Predicted probability of high risk based on age intervals.
+The median predicted probability of high risk is shown as a decimal, where age intervals [20, 30) and [50, 60) have relatively higher predicted probabilities of high risk.
 #| echo: false
@@
 
-## Interpreting High Risk Predictions {#sec-firstpoint}
+## Interpreting Results {#sec-firstpoint}
 
@@
 The logistic regression model, designed to predict the probability of high risk, effectively highlighted the relative importance of age, sex, and region in shaping these probabilities.
-Notably, rural regions consistently showed higher predicted probabilities of high risk compared to urban regions.
-This disparity underscores the challenges faced by rural areas, such as limited access to healthcare infrastructure, which may contribute to higher case fatality rates (CFR).
 
-Similarly, the analysis revealed that males are more likely to be classified as high risk than females, aligning with previous research suggesting that biological differences, comorbidities, and health-seeking behaviors could account for this discrepancy.
-Age also played a pivotal role, with certain age intervals, such as 10–20 and 30–40, showing significantly higher probabilities of high risk.
+### Predicted Probability of High Risk by Region {#sec-discussion_region}
+
+Based on @fig-model_region, there is a higher density of individuals predicted to have high risk in the urban regions than in the rural regions.
+This disparity underscores the potential challenges faced by urban regions, such as overcrowding and not enough healthcare, which may contribute to higher case fatality rates (CFR).
+Since patients of COVID-19 were encouraged to stay at home even after testing positive at a hospital, patients who lived in urban regions may not have been able to access quick enough healthcare if their symptoms were to worsen rapidly.
+This could be reflected in the higher predicted risks for these communities.
+Overall, the predicted probability of high risk between the two regions were not much different, potentially reflecting that both regions experience certain challenges in healthcare and are similar in the amount and timeliness of healthcare received.
+
+Another aspect of the figures that is worth pointing out, if not a more important aspect, is the overall shape.
+There are significantly higher densities at the extremes of the predicted probability of high risk than moderate levels for both urban and rural regions.
+A potential explanation of this is that individuals have high disparities in access to healthcare regardless of the region they live in.
+In other words, people either had really good, timely healthcare or really bad healthcare in both urban and rural regions.
+This indicates that the region an individual lived in may not have been the strongest predictor the high risk.
+More research and analysis are needed to identify the underlying variables that correlate with this pattern.
+
+### Predicted Probability of High Risk by Sex {#sec-discussion_sex}
+
+The analysis revealed that there are no significant differences in predicted probability of high risk in males and females.
+The size of the difference is extremely small, especially compared to the IQR highlighted in @fig-model_sex.
+The potential differences can be attributed to biological differences, comorbidities, and health-seeking behaviors.
+There may be a slight difference between sexes in the sensitivity to one's own health conditions, willingness to seek healthcare, and other aspects.
+However, the analysis shows that the high risk is essentially equal for the two sexes.
+There is also an extremely large distribution of predicted probabilities within sexes as shown by the IQR.
+Therefore, more research and analysis will be needed to conclude that there are or are no sexual differences in the predicted probability of high risk.
+
+### Predicted Probability of High Risk by Age {#sec-discussion_age}
+
+Age also played a pivotal role, with certain age intervals, such as 20-30 and 50-60, showing significantly higher probabilities of high risk.
+There is no clear trend where higher or lower age groups have increased or decreased predicted risk.
 This pattern may reflect complex interactions between behavioral, immunological, and societal factors influencing COVID-19 outcomes.
+For example, 20-30 year-olds may be more engaged in social activities and more constantly in the workplace.
+They may also be less likely to have started families with children and be less careful with their health and hygiene.
+50-60 year-olds may still be out in the work field with relatively high social interactions while also have reduced or decreasing immunity to pathogens.
 Understanding these relationships is important for tailoring public health interventions and resource allocation strategies.
+Further research with more targeted variables and measures such as frequency of social activity and overall health conditions may be necessary to find key relationships between age and probability of high risk.
+
+## Interpreting Predicted Probability of High Risk for COVID-19 {#sec-highrisk}
+
+Based on the model set-up [@sec-model_setup], the probability of high risk is the predicted probability that the calculated level of high risk based on demographic information of a specific datapoint is above the median level of high risk for the dataset.
+In terms of the real-world, this is the chance that someone is more likely to die from COVID-19 compared to the rest of the population.
+
+## High Risk Mechanisms in Public Health Research {#sec-publichealth}
+
+The analysis of high risk is commonly used in public health settings, especially in epidemiology.
+The high risk approach allows the analysis to better identify the groups that are at higher risk or more likely to yield a negative outcome from the disease [@platt].
+In the case of this analysis, the high risk groups that are identified are those who are more likely to find COVID-19 fatal rather than mild.
+
+The results generally show moderate to weak correlations and relationships between the different demographic characteristics and the predicted probability of high risk.
+This is common depending on the topic of focus.
+Since COVID-19 is being examined, this situation is considered normal since COVID-19 does not have obvious strong predictors [@marin].
+
+Returning to the discussion of the use of the high risk approach in public health research, it is an increasingly popular approach due to the strong focus it has on individuals and communities that are more at risk [@platt].
+This can direct actions and prevention to those who are most in need and can be especially important when the disease is fatal, such as COVID-19.
+Although this analysis shows little difference between different demographic characteristics, the approach can still have an effect when large populations are considered.
+As for other cases where there are clear differences in risk, for example, using age as a predictor of cancer and implementing closer monitoring for older individuals, may lead to more lives saved [@bastian].
 
@@
 
-The findings of this study have implications for future pandemic preparedness and response.
+A key takeaway from this study is the intersection between demographic factors and social determinants of health.
+There are specific demographic groups that may have well-documented disadvantages in accessing healthcare, which may be reflected as higher risk.
+Addressing these social determinants, such as income inequality, education, and healthcare infrastructure, could have long-term benefits for public health beyond the current pandemic.
+Interventions should not only focus on immediate health needs but also work toward alleviating these systemic disparities.
+
 By identifying demographic and regional factors associated with higher probabilities of high risk, policymakers can develop targeted interventions to mitigate disparities.
@@
 
-Thirdly, the data obtained for COVID-19 is obviously in the past and no longer an issue that is of primary concern.
+The patterns identified do not show strong correlations.
+The relationships between the key predictors (age, sex, and region) may not fully capture the complexity of the underlying risk factors.
+There could be additional variables that influence risk but were not included in this analysis.
+Furthermore, the correlations between these predictors and the high-risk classification may be influenced by other unmeasured or unobserved factors.
+These aspects of the dataset require further exploration in future research to better understand the full range of factors that contribute to COVID-19 risk and to refine the models used for risk classification.
+
+Finally, the data obtained for COVID-19 is obviously in the past and no longer an issue that is of primary concern.
 This issue is in line with the generalizability issue due to limited surveying countries.
 The patterns and models may be limited in generalizability due to the sole focus on COVID-19.
+The data itself also has limitations of not fully representing the population due to data collection methods.
+Only groups of individuals who report to hospitals or governments were sampled, leading to potential biases.
 
@@
 
-## Appendix A: Survey and Sampling Details {#sec-appendixa}
+## Appendix A: Raw Data {#sec-appendixraw}
+
+```{r}
+#| label: fig-head_analysis_data
+#| fig-cap: First six rows of the dataset after cleaning. Shows columns including key variables including region, date, sex, age, cases, deaths, and CFR.
+#| echo: false
+#| warning: false
+#| message: false
+
+# Load the necessary library
+library(knitr)
+
+# Display the first few rows of analysis_data
+head_table <- head(analysis_data)
+
+# Render the table with kable
+kable(
+  head_table,
+  booktabs = TRUE,  # Add better table styling
+)
+```
+
+## Appendix B: Survey and Sampling Details {#sec-appendixa}
 
@@
 Data collection follows a systematic approach.
-For primary sources, COVerAGE-DB retrieves data directly from official government dashboards, public health reports, and institutional publications (@cite-dowd2020).
+For primary sources, COVerAGE-DB retrieves data directly from official government dashboards, public health reports, and institutional publications [@dowd2020].
 This ensures that the dataset reflects real-world phenomena with minimal distortion.
@@
 The data is observational, collected without experimental controls.
-As a result, confounding factors such as healthcare system differences, socioeconomic disparities, and public health policies must be carefully considered when interpreting results (@cite-ioannidis2021).
+As a result, confounding factors such as healthcare system differences, socioeconomic disparities, and public health policies must be carefully considered when interpreting results [@ioannidis2021].
 For instance, differences in testing rates across regions can influence the apparent number of cases and consequently impact measures like CFR.
 
-## Appendix B: Model Selection Details {#sec-appendixb}
+## Appendix C: Idealized Methodology {#sec-appendix_method}
+
+The data collection method currently used in the dataset has some key limitations of not being able to reach the entire population.
+The idealized methodology will address these issues by implementing a new way of data collection where, if done in 2022, will provide a more representative dataset of COVID-19 data and information.
+
+The idealized methodology would have in-person PCR testing for the sample to gather data on the number of tests, number of cases, age, sex, and region.
+The number of deaths can be difficult to survey in-person without the support of government agencies.
+Therefore, the data for number of deaths can still be collected from relevant government institutions, as Canadians are required to report deaths to the government.
+This combined methodology will allow a more accurate dataset of COVID-19 and demographics information.
+
+### Population
+
+The population, given that the goal of the analysis is to generalize to the Canadian population, would be people living in Canada permanently.
+This includes all Canadian citizens, permanent residence, and any individuals with a Visa that extends for more than one year (2022).
+
+### Frame
+
+In the case of this methodology, when there are no financial limitations in play, the frame would be the same as the population.
+Since there are no restrictions on what households or Canadian citizens can't be sampled, the frame is essentially a subset of the population that is the entire population.
+
+### Sample and Sampling Method
+
+The sample would be a subset from the population.
+The idealized method must ensure that the sample consists of a representative group of individuals.
+The method to be used will be systematic sampling, where people would be hired to visit 1 of every 10 houses or apartments (or other units of living) and sample everyone within that household.
+This will avoid any sampling biases such as only sampling people who go out (if sampled on the streets) or only sampling people who drive (if sampling on roads).
+
+For every household sampled, the surveyor will conduct a PCR test for each person in the household, record each test, record the number of tests that return as positive as the number of cases, and gather demographic information of each person in the household such as such as sex and age.
+Region of the household should already be known and should also be recorded as part of the dataset.
+
+This should result in a dataset with a sufficient sample size.
+Given that there is an estimated 15.3 million households in 2022 [@households], the sample will consist of 1,530,000 households.
+There are an estimated 3 people per household [@familysize], leaving a sample of around 4.59 million people.
+
+The systematic sampling method ensures that a sufficient sample size will be gathered and allows rather accurate prediction of the number of datapoints within the sample.
+It also ensures for uniform coverage and that sampling wouldn't be biased towards certain communities or subgroups.
+However, limitations could surface if the selected samples are not responding, not available, or refuse to co-operate.
+The systematic sampling method makes it hard to change the sample size if sitations like this were to occur.
+
+The sampling procedure will be repeated monthly until the end of the year to gather time-related data.
+
+### Handling Non-response
+
+Non-response could mean not being at home during the visit of the surveyor or not willing to participate.
+The initial response to this could be to re-visit another time and attempt to persuade the participant to co-operate.
+
+If a small number of households refuse to participate after repeated visits, the broad demographic information (such as region) can be recorded and later adjusted for during the data anlysis process.
+However, if a large number of households refuse to participate, specific parts of the data may need to be re-sampled or dropped, which may cause biases in the analysis.
+An attempt to address this may be to adjust the dataset based on known demographic information, but certain biases may be unavoidable.
+
+### Idealized Procedure
+
+**Consent:** The surveyor will read this to the participant of the test before conducting the tests and ask for their signature.
+
+"Thank you for your will in potentially co-operating in this COVID-19 study.
+By participating in this study, you agree to provide demographic information and respond to questions related to COVID-19.
+The procedure will require conducting PCR test for you and asking you basic questions regarding your demographic characteristics.
+Your participation is voluntary, and all data will be kept confidential and anonymous and used solely for research purposes.
+You may withdraw from the study at any time without penalty.
+If you have any questions about the study, please feel free to ask.
+If you acknowledge your rights and are willing to consent to this study, please sign."
+
+**PCR:** The surveyor will conduct a PCR test for individuals who have consented to the study.
+
+**COVID-19 Information:** The surveyor will record the number of PCR tests conducted in the household and the number of positives.
+
+**Demographic Questions:** The surveyor will ask the following questions or record the known answer. \begin{itemize}
+  \item What is your biological sex at birth?
+  \item What is your age?
+  \item What area do you live in? Urban or rural?
+\end{itemize}
+
+**Debrief:** The surveyor will read the following to the participants after conducting the test and asking the questions.
+
+"Thank you for participating in this study.
+The information you provided on demographic factors and the test done for COVID-19 positivity will help us better understand the impact of the pandemic across different demographic groups.
+Your responses will contribute to valuable research aimed at improving public health strategies.
+If you have any questions or would like further information about the study, please feel free to contact us at "brucejc.zhang\@mail.utoronto.ca"
+
+**Data Consolidation:** The data collected by surveyors will be matched with the death-data from governments and hospitals at the end of the year of 2022 and will prepare the data for analysis for any patterns or trends as done in this analysis.
+
+## Appendix D: Model Selection Details {#sec-appendixb}
 
@@
 #| label: fig-appendixb
-#| fig-cap: Linear Regression Model for dataset
+#| fig-cap: Linear Regression Model for dataset. No clear pattern can be observed. Weak predictive power is seen. 
 #| echo: false
@@
 
-## Appendix C: Model Validation Details {#sec-appendixc}
+## Appendix E: Model Validation Details {#sec-appendixc}
 
@@
 #| label: fig-appendix_roc
-#| fig-cap: ROC Curve and AUC Value for Logistic Regression Model
+#| fig-cap: ROC Curve and AUC Value for Logistic Regression Model.
+The ROC curve and AUC values show that the logistic regression model yields a moderately high predictive performance for the dataset.
 #| echo: false
@@
 library(ggplot2)
+library(here)
+library(arrow)
 set.seed(21)
 
-analysis_data <- read.csv(here("data/02-analysis_data/modeling_data.csv"))
+analysis_data <- read_parquet(here("data", "03-model_data", "modeling_data.parquet"))
 

### 2024-12-03 scripts/00-simulate_data.R

@@ -7,9 +7,9 @@
 # Pre-requisites: The `tidyverse` package must be installed
 # Any other information needed? Make sure you are in the `COVID_demographics` rproj
 
-
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 set.seed(21)
 
 # Define parameters
@@ -41,4 +41,4 @@ simulated_data$Tests <- round(runif(nrow(simulated_data), min = 100, max = 50000
 head(simulated_data)
 
 #### Save data ####
-write_csv(simulated_data, "data/00-simulated_data/simulated_data.csv")
+write_parquet(simulated_data, "data/00-simulated_data/simulated_data.parquet")

### 2024-12-03 scripts/01-test_simulated_data.R

@@ -14,65 +14,89 @@
 #### Workspace setup ####
 library(tidyverse)
 library(dplyr)
+library(arrow)
+library(testthat)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+# Read the data
+analysis_data <- read_parquet("data/00-simulated_data/simulated_data.parquet")
 
-# Load necessary library
-library(dplyr)
+#### Tests with `testthat` ####
 
 # Test 1: Verify structure of the dataset
-str(simulated_data)
+test_that("Dataset has the expected structure", {
+  expect_s3_class(analysis_data, "data.frame")  # Check that the dataset is a data frame
+  expect_true(all(c("Country", "Region", "Code", "Date", "Sex", "Age", "AgeInt", "Cases", "Deaths", "Tests") %in% colnames(analysis_data)))
+})
 
 # Test 2: Check for missing values in key columns
-missing_summary <- simulated_data %>%
-  summarise(
-    Missing_Country = sum(is.na(Country)),
-    Missing_Region = sum(is.na(Region)),
-    Missing_Code = sum(is.na(Code)),
-    Missing_Date = sum(is.na(Date)),
-    Missing_Sex = sum(is.na(Sex)),
-    Missing_Age = sum(is.na(Age)),
-    Missing_Cases = sum(is.na(Cases)),
-    Missing_Deaths = sum(is.na(Deaths)),
-    Missing_Tests = sum(is.na(Tests))
-  )
-print("Missing values summary:")
-print(missing_summary)
+test_that("There are no missing values in key columns", {
+  missing_summary <- analysis_data %>%
+    summarise(
+      Missing_Country = sum(is.na(Country)),
+      Missing_Region = sum(is.na(Region)),
+      Missing_Code = sum(is.na(Code)),
+      Missing_Date = sum(is.na(Date)),
+      Missing_Sex = sum(is.na(Sex)),
+      Missing_Age = sum(is.na(Age)),
+      Missing_Cases = sum(is.na(Cases)),
+      Missing_Deaths = sum(is.na(Deaths)),
+      Missing_Tests = sum(is.na(Tests))
+    )
+  expect_equal(missing_summary$Missing_Country, 0)
+  expect_equal(missing_summary$Missing_Region, 0)
+  expect_equal(missing_summary$Missing_Code, 0)
+  expect_equal(missing_summary$Missing_Date, 0)
[…]

### 2024-12-03 scripts/03-clean_data.R

@@ -7,11 +7,11 @@
 # Pre-requisites: 
 # Any other information needed? Make sure you are in the `COVID_demographics` rproj
 
-install.packages("tidyr")
 # Load necessary libraries
 library(dplyr)
 library(tidyr)
 library(tidyverse)
+library(arrow)
 
 # Step 1: Remove unnecessary columns (if applicable)
 data_cleaned <- simulated_data %>%
@@ -72,4 +72,4 @@ data_cleaned <- data_cleaned %>%
 summary(data_cleaned)
 
 #### Save data ####
-analysis_data <- write_csv(data_cleaned, "data/02-analysis_data/analysis_data.csv")
+analysis_data <- write_parquet(data_cleaned, "data/02-analysis_data/analysis_data.parquet")

### 2024-12-03 scripts/04-test_analysis_data.R

@@ -9,112 +9,100 @@
 # Any other information needed? Make sure you are in the `COVID_demographics` rproj
 
 
-# Load necessary libraries
+#### Workspace setup ####
+library(tidyverse)
 library(dplyr)
-library(ggplot2)
-
-# Test the integrity of the analysis dataset
-message("Testing the analysis data...")
+library(arrow)
+library(testthat)  # For unit testing
 
 # Step 1: Load the data
-# Replace 'analysis_data_path' with the actual path to your data file
-analysis_data_path <- "data/02-analysis_data/analysis_data.csv"
-analysis_data <- read.csv(analysis_data_path)
+analysis_data_path <- "data/02-analysis_data/analysis_data.parquet"
+analysis_data <- read_parquet(analysis_data_path)
 
-# Step 2: Basic Structure and Preview
-message("Previewing the dataset...")
-print(head(analysis_data))
-print(str(analysis_data))
+#### Tests with `testthat` ####
 
-# Step 3: Check for Missing Values
-message("Checking for missing values...")
-missing_summary <- analysis_data %>%
-  summarise(across(everything(), ~ sum(is.na(.)))) %>%
-  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Missing_Count")
+# Test 1: Check the basic structure of the dataset
+test_that("Dataset has the expected structure", {
+  expect_s3_class(analysis_data, "data.frame")  # Dataset should be a data frame
+  expect_true(all(c("Country", "Region", "Code", "Date", "Sex", "Age", "AgeInt", "Cases", "Deaths", "Tests") %in% colnames(analysis_data)))
+})
 
-print(missing_summary)
+# Test 2: Check for missing values in key columns
+test_that("There are no missing values in key columns", {
+  missing_summary <- analysis_data %>%
+    summarise(across(everything(), ~ sum(is.na(.)))) %>%
+    pivot_longer(cols = everything(), names_to = "Variable", values_to = "Missing_Count")
+  
+  expect_equal(missing_summary$Missing_Count, rep(0, length(missing_summary$Missing_Count)))  # Ensure no missing values in any column
+})
 
-# Step 4: Check for Duplicates
-message("Checking for duplicate rows...")
-duplicates <- analysis_data %>%
-  duplicated() %>%
-  sum()
-message(paste("Number of duplicate rows:", duplicates))
[…]

### 2024-12-03 scripts/05-exploratory_data_analysis.R

@@ -1,37 +1,65 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models the exploratory data analysis of the COVID-19 dataset
+# Author: Bruce Zhang
+# Date: 3 December 2024
+# Contact: brucejc.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+  # - 03-clean_data.R must have been run
+# Any other information needed? 
 
-
-#### Workspace setup ####
+# Load necessary libraries
 library(tidyverse)
-library(rstanarm)
+library(dplyr)
+library(ggplot2)
+library(corrplot) 
+library(DataExplorer) 
 
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+# Set random seed for reproducibility
+set.seed(21)
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# Step 1: Load the data (make sure to change the file path to your actual dataset)
+analysis_data <- read_parquet(here("data", "02-analysis_data", "analysis_data.parquet"))
 
+# Step 2: Data Overview
+# View the structure of the data
+str(analysis_data)
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
+# Check for missing values
+summary(analysis_data)
 
+# Get the summary statistics of numerical columns
+summary_stats <- analysis_data %>%
+  select_if(is.numeric) %>%
+  summary()
 
+# Step 3: Exploratory Visualizations
+# Histogram for distribution of numerical variables (e.g., Cases, Deaths, Tests)
+ggplot(analysis_data, aes(x = Cases)) +
[…]

### 2024-12-03 scripts/06-model_data.R

@@ -15,6 +15,7 @@ library(ggplot2)
 library(dplyr)
 library(caret)
 library(pROC)
+library(arrow)
 
 
 # Building a logistic regression model for predicted high risk and age
@@ -214,5 +215,5 @@ ggplot(filtered_data, aes(x = Region, y = Predicted_Probability, fill = Region))
 # Step 6: Save the Model for Reproducibility
 saveRDS(logistic_model, file = "models/logistic_model_region_high_risk.rds")
 
-write.csv(analysis_data, "data/02-analysis_data/modeling_data.csv", row.names = FALSE)
+write_parquet(analysis_data, "data/03-model_data/modeling_data.parquet")
 

### 2024-12-03 scripts/01-test_simulated_data.R
Updated scripts comments

@@ -20,8 +20,6 @@ library(testthat)
 # Read the data
 analysis_data <- read_parquet("data/00-simulated_data/simulated_data.parquet")
 
-#### Tests with `testthat` ####
-
 # Test 1: Verify structure of the dataset
 test_that("Dataset has the expected structure", {
   expect_s3_class(analysis_data, "data.frame")  # Check that the dataset is a data frame

### 2024-12-03 scripts/02-download_data.R

@@ -49,10 +49,6 @@ cat(
 
 
 #### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-Data <- read.csv(output_file)
-
 write_csv(Data, "data/01-raw_data/raw_data.csv") 
 
          

### 2024-12-03 scripts/03-clean_data.R

@@ -18,7 +18,7 @@ data_cleaned <- simulated_data %>%
   select(Country, Region, Date, Sex, Age, Cases, Deaths, Tests)
 
 # Step 2: Handle missing values
-# Replace missing numeric values with 0 (if appropriate)
+# Replace missing numeric values with 0
 data_cleaned <- data_cleaned %>%
   mutate(
     Cases = ifelse(is.na(Cases), 0, Cases),
@@ -45,14 +45,13 @@ data_cleaned <- data_cleaned %>%
     Deaths = as.numeric(Deaths),
   )
 
-# Remove duplicate rows (if any)
+# Step 5: Remove duplicate rows
 data_cleaned <- data_cleaned %>%
   distinct() %>%
   filter(Country == "Canada") %>%
   filter(Region %in% c("Urban", "Rural"), Sex %in% c("m", "f"))
 
-# Check for and handle outliers (optional)
-# For example, cap unrealistic values for 'Cases', 'Deaths', or 'Tests'
+# Step 6: Check for and handle outliers
 data_cleaned <- data_cleaned %>%
   mutate(
     Cases = ifelse(Cases > 100000, 100000, Cases),
@@ -63,8 +62,8 @@ data_cleaned <- data_cleaned %>%
 data_cleaned <- data_cleaned %>%
   filter(Deaths <= Cases)
 
-# Add calculated columns (if needed)
-# Example: Case Fatality Rate (CFR)
+# Step 7: Add calculated columns
+# Case Fatality Rate (CFR)
 data_cleaned <- data_cleaned %>%
   mutate(CFR = ifelse(Cases > 0, (Deaths / Cases) * 100, NA))
 

### 2024-12-03 scripts/04-test_analysis_data.R

@@ -15,11 +15,9 @@ library(dplyr)
 library(arrow)
 library(testthat)  # For unit testing
 
-# Step 1: Load the data
+# Load the data
 analysis_data_path <- "data/02-analysis_data/analysis_data.parquet"
 analysis_data <- read_parquet(analysis_data_path)
-
-#### Tests with `testthat` ####
 
 # Test 1: Check the basic structure of the dataset
 test_that("Dataset has the expected structure", {

### 2024-12-03 scripts/05-exploratory_data_analysis.R

@@ -14,11 +14,9 @@ library(dplyr)
 library(ggplot2)
 library(corrplot) 
 library(DataExplorer) 
-
-# Set random seed for reproducibility
 set.seed(21)
 
-# Step 1: Load the data (make sure to change the file path to your actual dataset)
+# Step 1: Load the data
 analysis_data <- read_parquet(here("data", "02-analysis_data", "analysis_data.parquet"))
 
 # Step 2: Data Overview
@@ -34,7 +32,7 @@ summary_stats <- analysis_data %>%
   summary()
 
 # Step 3: Exploratory Visualizations
-# Histogram for distribution of numerical variables (e.g., Cases, Deaths, Tests)
+# Histogram for distribution of numerical variables
 ggplot(analysis_data, aes(x = Cases)) +
   geom_histogram(bins = 30, fill = "steelblue", color = "black", alpha = 0.7) +
   labs(title = "Distribution of Cases", x = "Cases", y = "Frequency")

### 2024-12-03 scripts/06-model_data.R

@@ -8,9 +8,7 @@
   # - 03-clean_data.R must have been run
 # Any other information needed? Make sure you are in the `COVID_demographics` rproj
 
-# Install and load necessary pacakges/libraries
-install.packages("caret")
-install.packages("pROC")
+# Load necessary pacakges/libraries
 library(ggplot2)
 library(dplyr)
 library(caret)

### 2024-12-03 scripts/07-replications.R

@@ -1,19 +1,213 @@
 #### Preamble ####
-# Purpose: Replicated graphs from... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Validates the models constructed.
+# Author: Bruce Zhang
+# Date: 3 December 2024
+# Contact: brucejc.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+  # - 03-clean_data.R must have been run
+# Any other information needed? 
 
 
-#### Workspace setup ####
-library(tidyverse)
-# [...UPDATE THIS...]
+# Load necessary pacakges/libraries
+library(ggplot2)
+library(dplyr)
+library(caret)
+library(pROC)
+library(arrow)
 
-#### Load data ####
-# [...ADD CODE HERE...]
+
+# Building a logistic regression model for predicted high risk and age
+# Step 0: Define High_Risk variable in the dataset
+analysis_data <- analysis_data %>%
+  mutate(
+    Death_Rate = Deaths / (Cases + 1e-6), # Calculate death rate (avoid division by zero)
+    Log_CFR = log(CFR + 1),
+    High_Risk = ifelse(Death_Rate > median(Death_Rate, na.rm = TRUE), 1, 0) # Classify high-risk regions
+  )
+
+# Step 1: Build the logistic regression model
+logistic_model <- glm(
+  High_Risk ~ Age + Region + Sex + Cases, # Replace with relevant predictors
+  data = analysis_data,
+  family = binomial()
+)
+
+# Step 2: Model Summary and Diagnostics
+cat("Model Summary:\n")
+print(summary(logistic_model))
+cat("\nAIC:", AIC(logistic_model), "\nBIC:", BIC(logistic_model), "\n")
+
+# Step 3: Predicted Probabilities and Classification
+analysis_data <- analysis_data %>%
+  mutate(
+    Predicted_Probability = predict(logistic_model, type = "response"),
+    Predicted_Class = ifelse(Predicted_Probability > 0.5, 1, 0)  # Classification threshold
+  )
+
+# Step 4: Confusion Matrix
[…]

### 2024-12-03 paper/paper.qmd
Updated formating and references

@@
 This analysis looks at COVID-19 data from 2022 in Canada and analyzes how the different population and individual-level demographic variables, such as age, sex, and region of life, influenced the risk of fatality from COVID-19.
+The analysis focuses on isolating these demographic factors and examining their individual effects.
 The paper uses the demographic data gathered from 2022 to predict the characteristics of populations and individuals that may be at higher risk of fatality once they have been diagnosed with COVID.
@@
 I use the statistical programming language R [@citeR] to analyze the data and to create graphs and models.
-Th packages that were used include tidyverse [@tidyverse], tidyr [@tidyr], dplyr [@dplyr], arrow [@arrow], DataExplorer [@dataexplorer], corrplot [@corrplot], caret [@caret], and pROC [@pROC].
+Th packages that were used include tidyverse [@tidyverse], tidyr [@tidyr], dplyr [@dplyr], arrow [@arrow], DataExplorer [@dataexplorer], corrplot [@corrplot], caret [@caret], httr [@httr], here[@here], and pROC [@pROC].
 My data [@covid_data] was obtained from COVerAGE-DB, which was housed in Open Science Framework.
@@
 #| fig-cap: Predicted probability of high risk based on age intervals.
-The median predicted probability of high risk is shown as a decimal, where age intervals [20, 30) and [50, 60) have relatively higher predicted probabilities of high risk.
+The median predicted probability of high risk is shown as a decimal, where age intervals 20-30 and 50-60 have relatively higher predicted probabilities of high risk.
 #| echo: false
@@
 ```
-
+\newpage
 ### Odds Ratio with Confidence Intervals

### 2024-12-03 paper/paper.qmd
Updated prose

@@
 Another key advantage of the logistic regression model in this context is that it provides interpretable coefficients, which allow for an understanding of how specific predictors influence the odds of being high risk.
-This interpretability is crucial for public health applications.
+This interpretability is very important for public health applications.
 Furthermore, logistic regression is computationally efficient, making it ideal for datasets with relatively straightforward structures like this one.
@@
 This disparity underscores the potential challenges faced by urban regions, such as overcrowding and not enough healthcare, which may contribute to higher case fatality rates (CFR).
-Since patients of COVID-19 were encouraged to stay at home even after testing positive at a hospital, patients who lived in urban regions may not have been able to access quick enough healthcare if their symptoms were to worsen rapidly.
+Since patients of COVID-19 were encouraged to stay at home even after testing positive at a hospital, patients who lived in urban regions may not have been able to access quick enough healthcare if their symptoms were to worsen quickly.
 This could be reflected in the higher predicted risks for these communities.
@@
 
-The analysis revealed that there are no significant differences in predicted probability of high risk in males and females.
+The analysis showed no significant differences in predicted probability of high risk in males and females.
 The size of the difference is extremely small, especially compared to the IQR highlighted in @fig-model_sex.
@@
 Another limitation is the reliance on aggregate data, which may obscure individual-level variability and interactions.
-For example, the interplay between age, sex, and comorbidities at the individual level could yield more nuanced insights.
+For example, the interplay between age, sex, and comorbidities at the individual level could yield more relevant information.
 Additionally, the classification of rural and urban regions may oversimplify the diversity within these categories, such as the distinction between suburban and remote rural areas.
@@
 Furthermore, the correlations between these predictors and the high-risk classification may be influenced by other unmeasured or unobserved factors.
-These aspects of the dataset require further exploration in future research to better understand the full range of factors that contribute to COVID-19 risk and to refine the models used for risk classification.
+These aspects of the dataset require further research to better understand the full range of factors that contribute to COVID-19 risk and to refine the models used for risk classification.
 
@@
 The database also ensures quality Control through running automated checks and manual reviews to detect inconsistencies, missing values, and potential errors in data entry.
-Such measures are crucial given the heterogeneity in reporting standards across countries and regions.
+Such measures are key given the heterogeneity in reporting standards across countries and regions.
 
@@
 The information you provided on demographic factors and the test done for COVID-19 positivity will help us better understand the impact of the pandemic across different demographic groups.
-Your responses will contribute to valuable research aimed at improving public health strategies.
+Your responses will contribute to research aimed at improving public health strategies.
 If you have any questions or would like further information about the study, please feel free to contact us at "brucejc.zhang\@mail.utoronto.ca"

### 2024-12-03 paper/paper.qmd
Updated formatting

@@
 ---
-title: "Number of COVID-19 Cases in Canada in 2022 By Region, Age, and Sex"
+title: "COVID-19 Cases and Deaths in Canada in 2022 By Region, Age, and Sex"
 subtitle: "Predicting the Factors that Influence the Probability of High Risk of COVID-19"
@@
 Rural).
-This is a density plot showing that the extremes of predicted probability of high risk are more frequent for both urban and rural regions.
+The density of probabilities of high risk shows that the extremes of predicted probability of high risk are more frequent for both urban and rural regions.
 #| echo: false
@@
 #| label: fig-model_sex
-#| fig-cap: Predicted probability of high risk based on sex.
-The box-and-whiskers plots show that the median value of the predicted probability of high risk is similar for males and females.
-There is a large IQR for both.
+#| fig-cap: Predicted probability of high risk based on sex. The median value of the predicted probability of high risk is shown to be similar for males and females. There is a large IQR for both. 
 #| echo: false

### 2024-12-14 paper/paper.qmd
Updated formatting
Removed typos

@@
 
-The data that was collected through hopistals and government institutions was gathered by COVerAGE-DB, a database that focuses on COVID related data.
+The data that was collected through hospitals and government institutions was gathered by COVerAGE-DB, a database that focuses on COVID related data.
 COVerAGE-DB consolidates this information into the dataset that was used in the analysis.
@@
 For this particular dataset, the key predictor variables are all categorical, even age is categorized by intervals.
-Therefore, it was extremely necessary for the model to be able to handl categorical data, which the logistic regression model is capable of.
+Therefore, it was extremely necessary for the model to be able to handle categorical data, which the logistic regression model is capable of.
 
@@
 However, limitations could surface if the selected samples are not responding, not available, or refuse to co-operate.
-The systematic sampling method makes it hard to change the sample size if sitations like this were to occur.
+The systematic sampling method makes it hard to change the sample size if situations like this were to occur.
 
@@
 
-If a small number of households refuse to participate after repeated visits, the broad demographic information (such as region) can be recorded and later adjusted for during the data anlysis process.
+If a small number of households refuse to participate after repeated visits, the broad demographic information (such as region) can be recorded and later adjusted for during the data analysis process.
 However, if a large number of households refuse to participate, specific parts of the data may need to be re-sampled or dropped, which may cause biases in the analysis.
@@
 ```
-
+\newpage
 ## Appendix E: Model Validation Details {#sec-appendixc}
@@
 
+\newpage
 ### Confusion Matrix
